---
title: ""
output:
  html_document:
    theme: journal
df_print: paged
pdf_document: default
word_document: default
always_allow_html: yes
params:
  agency: defence.gov.au,dst.defence.gov.au
  agency_name: Department of Defence and Associated Agencies
  agency_aust_name: Department of Defence,Australian Signals Directorate

---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!exists("buyers")) {
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  library(httr)
  library(stringr)
  library(kableExtra)
  library(lubridate)
  library(googlesheets)
  source("../R/utilities.r")
  source("../R/plotly_functions.r")
  source("../R/assorted_functions.r")
  source("../R/scrape_austender.r")
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),rel_path_data(),latest_data_file()))
  #cns <- load_austender_cns() %>% post_process_cns()
}

```   
run date: `r date_to_string(Sys.Date())`  

```{r, echo = FALSE, message = FALSE}
ref_domain    <- agencies
email_domains <- tibble(
      email_domain = str_trim(unlist(str_split(params$agency,"\\,")))
                      ) %>% 
  left_join(ref_domain, by = "email_domain") %>% 
  transmute(Agency = agencyName,
            emails = email_domain)

```
## Agency summary report for <span style = "color:blue">`r params$agency_name`</span>      
<br />  

## Summary  
```{r, echo=FALSE}
b <- briefs %>%
  filter(email_domain %in% email_domains$emails) %>% 
  filter(status %in% c("closed", "live"))

b_summ <- b %>%
  summarise(
    total    = n(),
    earliest = min(published),
    latest   = max(published)
  )

c_agencies <- unlist(str_split(params$agency_aust_name, "\\,"))
c_summ <- contracts %>% 
  filter(Agency %in% c_agencies) %>% 
  group_by(sme_by_employees) %>% 
  summarise(
    count = n(),
    value = sum(Value),
    earliest = min(Publish.Date),
    latest   = max(Publish.Date)
  )

summary_table = tibble(
  name = c("total briefs published",
           "earliest brief",
           "latest brief",
           "total contracts reported",
           "contracts awarded to SMEs",
           "percentage awarded to SMEs",
           "total value of contracts reported",
           "value awarded to SMEs",
           "percentage of value awarded to SMEs",
           "earliest contract",
           "latest contract"
           ),
  values = c(b_summ$total, 
             date_to_string(b_summ$earliest), 
             date_to_string(b_summ$latest),
             sum(c_summ$count),
             c_summ[2,]$count,
             paste0(round(c_summ[2,]$count/sum(c_summ$count)*100),"%"),
             format_mil(sum(c_summ$value)),
             format_mil(sum(c_summ[2,]$value)),
             paste0(round(c_summ[2,]$value/sum(c_summ$value)*100),"%"),
             date_to_string(min(c_summ$earliest)),
             date_to_string(max(c_summ$latest)))
)

```
`r kable(summary_table, col.name = c("","")) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))`   


## Briefs published  
<br />
```{r, echo = FALSE, message = FALSE}
b <- briefs %>%
  filter(email_domain %in% email_domains$emails) %>% 
  filter(status %in% c("closed", "live"))

b_summ_day <- b %>% 
  group_by(published) %>% 
  count(published)

b_summ_month <- b %>% 
  mutate(month = floor_date(published, unit = "month")) %>% 
  count(month) %>% 
  right_join(datesSinceLaunch(unit = "month"), by = c("month" = "dates")) %>% 
  replace_na(list(n = 0))

p <- plot_ly(data = b_summ_day, x =~ published, y =~ n, type = "bar",
             name = "by day") %>% 
  add_trace(data = b_summ_month, x =~ month, y =~ n, type = "scatter",
            name = "by month", mode = "lines+markers") %>% 
  layout(title = "Briefs published by date",
         xaxis = list(title = ""),
         yaxis = list(title = "Number of briefs published"),
         legend = list(x = 0.05, y = 0.95))
```
`r nrow(b)` brief`r if (nrow(b) != 1) 's'` published, from 
`r date_to_string(min(b$published))` though to `r date_to_string(max(b$published))`.  
`r p`  

## Reported contracts by count   
<br />  
```{r, echo = FALSE}
c_counts <- contracts %>%
  filter(Agency %in% c_agencies) %>% 
  mutate(month = floor_date(Publish.Date, unit = "month")) %>% 
  group_by(month, sme_by_employees) %>% 
  summarise(
    count = n()
  ) %>% 
  spread(sme_by_employees, count) %>% 
  right_join(datesSinceLaunch(unit = "month"), by = c("month" = "dates")) %>% 
  rename(SME = `TRUE`,
         Enterprise = `FALSE`) %>% 
  replace_na(list(SME = 0, Enterprise = 0)) %>% 
  ungroup() %>% 
  mutate(
    SME_total        = cumsum(SME),
    Enterprise_total = cumsum(Enterprise),
    total            = SME_total + Enterprise_total,
    percent_sme      = round(SME_total / total * 100)
  ) %>% 
  filter(!is.nan(percent_sme))

p2 <- plot_ly(data = c_counts, x =~ month, y =~ SME, 
              type = "bar", name = "SME") %>% 
  add_bars(y =~ Enterprise, name = "Enterprise") %>% 
  add_trace(y =~ SME_total, name = "Total SME", type = "scatter", 
            mode = "lines+markers") %>% 
  add_trace(y =~ Enterprise_total, name = "Total enterprise", 
            type = "scatter", mode = "lines+markers") %>% 
  layout(barmode = "stack",
         title = "Count of contracts reported",
         xaxis = list(title = ""),
         yaxis = list(title = "count"),
         legend = list(x = 0.05, y = 0.95))
```
`r p2`   
  
```{r, echo = FALSE}
p3 <- plot_ly(data = c_counts, x =~ month, y =~ percent_sme, 
              type = "scatter", mode = "lines+markers") %>% 
  layout(title = "Percentage count of contracts awarded to SMEs",
         yaxis = list(title = "percent awarded to SME by count",
                      range = c(0,101)))
```
<br />  
`r p3`  

## Reported contracts by value  
```{r, echo = FALSE}  
c_values <- contracts %>% 
  filter(Agency %in% c_agencies) %>% 
  mutate(month = floor_date(Publish.Date, unit = "month")) %>% 
  group_by(month, sme_by_employees) %>% 
  summarise(
    value = sum(Value)
  ) %>% 
  spread(sme_by_employees, value) %>% 
  right_join(datesSinceLaunch(unit = "month"), by = c("month" = "dates")) %>% 
  rename(SME        = `TRUE`,
         Enterprise = `FALSE`) %>% 
  replace_na(list(SME = 0, Enterprise = 0)) %>% 
  ungroup() %>% 
  mutate(
    SME_total        = cumsum(SME),
    Enterprise_total = cumsum(Enterprise),
    total            = SME_total + Enterprise_total,
    percent_sme      = round(SME_total / total * 100)
  ) %>% 
  filter(!is.nan(percent_sme))
  
p4 <- plot_ly(data = c_values, x =~ month, y =~ SME_total, 
              type = "scatter", mode = "lines+markers", name = "SME") %>% 
  add_trace(y =~ Enterprise_total, name = "Enterprise") %>% 
  add_trace(y =~ total, name = "Total") %>% 
  layout(barmode = "stack",
         title = "Reported contract values",
         xaxis = list(title = ""),
         yaxis = list(title = "value"),
         legend = list(x = 0.05, y = 0.95))

```
`r p4`  
<br />

```{r, echo = FALSE}  
p5 <- plot_ly(data = c_values, x =~ month, y =~ percent_sme, 
              type = "scatter", mode = "lines+markers") %>% 
  layout(title = "Percentage value of contracts awarded to SMEs",
         yaxis = list(title = "percent awarded to SME by value",
                      range = c(0,101)))
```
`r p5`   

## Briefs published to date   
```{r, echo = FALSE}
b <- briefs %>%
  filter(email_domain %in% email_domains$emails) %>% 
  select(id, organisation, title, type, openTo, published) %>% 
  arrange(desc(published))
```
  
`r kables(b)`  

## Contracts reported to date  
```{r, echo = FALSE}
c <- contracts %>% 
    filter(Agency %in% c_agencies) %>% 
    select(Agency, CN.ID, Supplier.Name, 
           Supplier.ABN, Description, Publish.Date, 
           Value, sme_by_employees) %>% 
  arrange(desc(Publish.Date))
```
`r kables(c)`  
