---
title: "Dashboard Stats"
output:
  html_document:
  theme: cerulean
  toc: true
pdf_document: default
word_document: default
always_allow_html: yes
---
  
```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE,include=FALSE}
# if rendered in a new R session, load the session values
if (!exists("buyers")) {
  source("utilities.r")
  source("plotly_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  library(googlesheets)
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),"/../data/",substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
}

dsl   <- datesSinceLaunch()$dates
yms   <- unique(format(dsl,"%Y-%m"))[-1]
ymds  <- as.Date(paste0(yms,"-01")) - 1
```
## User satisfaction  
Calculated as the average monthly response from user satisfaction surveys on the Marketplace.  
User satisfaction scores are translated to 0-100 scale (Difficult=0,Ok=50,Easy=100)  
    
```{r,echo=FALSE}
f <- feedback %>% 
  mutate(month=format(date,"%Y-%m")) %>%
  filter(!is.na(cx)) %>%
  group_by(month) %>%
  summarise(responses=length(month),
            average_score=round(mean(cx*50)),
            #median=median(cx*50), # not interesting, as it will always be 0,50 or 100
            sd=round(sd(cx*50)))
plot_ly(f,x=~month,y=~average_score,type="scatter",mode="markers") %>%
  layout(title="User satisfaction",
         xaxis=list(title="Month"),
         yaxis=list(range=c(0,100),title="User satisfaction"))

```  
  
  
### Summary figures
`r kable(f)`  
  
## Cost per transaction  

```{r,echo=FALSE,include=F}
cprd      <- gs_title("Cost Per Transaction Data",verbose=F) %>% 
                gs_read(ws = "data",verbose=F) %>% 
                filter(!is.na(sessions))
```
`r kable(cprd)`

## Digital take-up  
Calculated as the percentage of commonwealth government organisations that have published a brief on the Digital Marketplace.  
  
Organisations included are 'principal entities' of type 'Non Corporate Commonwealth Entity' (NCCE) or 'Corporate Commonwealth Entity' (CCE)  
   
At 30 Sep 2017, there are 182 organisations.  

```{r,echo=FALSE}
br <- briefs %>% 
  filter(entities %in% c("NCCE","CCE")) %>%
  mutate(month=format(published,"%Y-%m"))
published_agencies <- sapply(ymds,
                             function(x,br) {
                               length(unique(br[br$published <= x,"email_domain"]))
                             },
                             br = br)
published_agencies <- round(published_agencies / 182 * 100)
plot_ly(x=ymds,y=published_agencies,type="scatter",mode="markers") %>%
    layout(title="Digital take-up",
         xaxis=list(title=""),
         yaxis=list(range=c(0,100),title="% of Commonwealth agencies"))
takeup <- data.frame(as_at_date = ymds,percent_takeup = published_agencies)
```  
  
`r kable(takeup)`  


## Completion rate  
Shows the percentage of published briefs that are reported in Austender. Only briefs published by reporting entities and have closed are included.  
```{r,echo=F}
br <- briefs %>% filter(reports=="yes",status=="closed")
briefs_by_date <- sapply(ymds,
                             function(x,br) {
                               dim(br[br$published <= x,])[1]
                             },
                             br = br)
contracts_by_date <- sapply(ymds,
                             function(x,co) {
                               dim(co[co$Publish.Date <= x,])[1]
                             },
                             co = contracts)
percs <- round(contracts_by_date / briefs_by_date * 100)[-1]
plot_ly(x=ymds[-1],y=percs,type="scatter",mode="markers") %>%
    layout(title="Completion rate",
         xaxis=list(title=""),
         yaxis=list(range=c(0,100),title="% of briefs reported as contracted"))
completion_rate <- data.frame(as_at_date=ymds[-1],percent_complete=percs)
```  
  
`r kable(completion_rate)`  
    
##  Total opportunities  
```{r,echo=F}
brief_count <- sapply(ymds,
             function(x,br) {
               dim(br[br$published <= x,])[1]
             },
             br = briefs)
brief_counts <- data.frame(as_at_date=ymds,published_briefs=brief_count) 
plot_ly(brief_counts,x=~as_at_date,y=~brief_count,type="scatter",mode="markers") %>%
    layout(title="Total opportunities",
         xaxis=list(title=""),
         yaxis=list(title="Opportunities published"))

```
  
`r kable(brief_counts)`  

## Registered buyers  

```{r,echo=F}
bu <- buyers %>% filter(active,!grepl("\\+",email_address))
buyer_count <- sapply(ymds,
                 function(x,bu) {
                   dim(bu[bu$created_at <= x,])[1]
                 },
                 bu = bu)
buyer_counts <- data.frame(as_at_date=ymds,registered_buyers=buyer_count) 
plot_ly(buyer_counts,x=~as_at_date,y=~buyer_count,type="scatter",mode="markers") %>%
    layout(title="Total registered buyers",
         xaxis=list(title=""),
         yaxis=list(title="Buyers"))

```  
  
`r kable(buyer_counts)`  
  
  
## Sellers  

```{r,echo=F}
seller_count <- sapply(ymds,
                      function(x,se) {
                        dim(se[se$joined <= x,])[1]
                      },
                      se = sellers)
seller_counts <- data.frame(as_at_date=ymds,registered_sellers=seller_count) 
plot_ly(seller_counts,x=~as_at_date,y=~seller_count,type="scatter",mode="markers") %>%
    layout(title="Total registered sellers",
         xaxis=list(title=""),
         yaxis=list(title="Sellers"))

```  
  
`r kable(seller_counts)`  
