---
title: "Digital Marketplace - 'Key Facts'"
output:
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,include=FALSE}
# if rendered in a new R session, load the session values
if (!exists("buyers")) {
  source("utilities.r")
  source("plotly_functions.r")
  source("assorted_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  library(lubridate)
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),"\\..",rel_path_data,substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
}

# filter inactive buyers
buyers <- buyers[buyers$active==TRUE,]
buyers <- buyers[!grepl("\\+",buyers$email_address),]
# need a better function to determine close date

```
**`r format(today(), format="%d %B %Y")`**  

## SME figures 
```{r, echo=FALSE}
no_panellists <- sum(sellers$dmp_framework&nchar(sellers$assessed_aoe) > 1)
no_sme        <- sum(sellers$dmp_framework&sellers$sme_by_employees&nchar(sellers$assessed_aoe) > 1)
perc_sme      <- round(no_sme/no_panellists*100,0)
tot_spend     <- round(sum(contracts$Value)/1000000,2)
c <- contracts %>% filter(contracts$sme_by_employees)
sme_spend     <- round(sum(c$Value)/1000000,2)
perc_sme_spend <- round(sme_spend/tot_spend*100,0)
tib           <- tibble('No. of panellists' = no_panellists,
                        'No. of SMEs'       = no_sme,
                        '% of SMEs'         = perc_sme,
                        'Total spend'       = tot_spend,
                        'Total SME spend'   = sme_spend,
                        '% of SME spend'    = perc_sme_spend)

```
`r kable(tib)`  

## Key Facts  
* The Digital Marketplace is an initiative under the National Innovation and Science
Agenda. Its purpose is to make is easier for smaller businesses to compete for the
governments spend, which was $6.2 billion in 2015-16, on information and
communications technology (ICT) product and services.  

* As at `r format(today(), format="%d %B %Y")`, $`r tot_spend`m of contracts have been reported to AusTender, with
`r perc_sme_spend` percent of that value awarded to SMEs.  

* As at `r format(today(), format="%d %B %Y")`, the Digital Marketplace has:  
    + `r dim(sellers)[1]` registered sellers,    
    + `r dim(buyers)[1]` registered buyers, and  
    + `r dim(briefs)[1]` opportunities posted.  

## Background:  
* The Digital Marketplace is a procurement platform that launched on 29 August 2016.   

* Initially, the Digital Marketplace utilised the Digital Services Professionals Panel (DSPP).
 The Digital Services Professionals Panel was a panel established by the DTA through an
approach to market published early in 2016, designed as an interim tool prior to the
Digital Marketplace coming online; however, the Digital Marketplace launched ahead of
schedule, so used the panel as its underlying procurement method.  

* In February 2017, the DTA went out to market again and published a request for panel to
AusTender for the Digital Marketplace Panel.  

* In February, the Marketplace also made it possible for sellers to apply online at any time.  

* The Digital Marketplace Panel remains open for new sellers to join, and the Marketplace
assess applications on an ongoing basis.  

* In April 2017 the Digital Marketplace transitioned to the Digital Marketplace Panel as its
underlying procurement method.  

```{r,echo=FALSE}
n_b       <- dim(briefs)[1]
n_ncce    <- sum(briefs$entities=="NCCE")
perc_cce <- 100 - round(n_ncce / n_b * 100,0)
```

* Additionally, `r perc_cce` percent of opportunities published to the Digital Marketplace were from
agencies that do not report to AusTender, such as corporate commonwealth entities and
local governments, so the figure of contracts awarded is likely a similar percentage higher.