---
title: "Monthly Reports Data Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
start <- as.Date("2017-09-11")
end   <- as.Date("2017-09-26")
fy_start <- as.Date("2017-07-01")
reference_file <- "..\\data\\2017-09-19.Rdata"

# if rendered in a new R session, load the session values
if (!exists("buyers")) {
  source("utilities.r")
  source("plotly_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  ## load the files - ensure these files are already produced
  #load(paste0(getwd(),"/../data/",substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
  load(reference_file)
}


```  
Date range for report:  
Starts `r start`  
Ends `r end`  
Current financial year: `r fy_start`  
  
## Registered Entities  
  
```{r,echo=FALSE}
agencies_counts <- buyers %>%
                group_by(email_domain,category) %>% 
                summarise(count=length(email_domain)) %>%
                group_by(category) %>%
                summarise(count = length(category))
agencies_counts$percent <- round(agencies_counts$count/sum(agencies_counts$count) * 100,digits=1)
agencies_counts <- rbind(agencies_counts,c("Total agencies",sum(agencies_counts$count),100.0))

```  

`r kable(agencies_counts)`     
  
# Top buyers in month  

```{r,echo=FALSE}
buyer_counts <- briefs %>%
  filter(published >= start,published <= end) %>%
  group_by(agencyName) %>%
  summarise(count=length(agencyName)) %>%
  arrange(desc(count))
```

`r kable(buyer_counts)`     
  
# How are we encouraging competition  

```{r,echo=FALSE}
# sme figures
#m_contracts <- left_join(contracts,sellers,by=c("Supplier.ABN" = "abn"))
m_contracts <- contracts

sme_emps  <- m_contracts %>% 
  filter(Publish.Date <= end) %>%
  group_by(sme_by_employees) %>% 
  summarise(value = sum(Value), count = length(sme_by_employees))
sme_emps$by_value <- sme_emps$value / sum(sme_emps$value) * 100
sme_emps$by_count <- sme_emps$count / sum(sme_emps$count) * 100
```

`r kable(sme_emps)`  
  
## Contract values  

```{r,echo=FALSE}
m_contracts <- contracts %>%
  filter(Publish.Date <= end)
total_value <- sum(m_contracts$Value)
total_contracts <- dim(m_contracts)[1]

m_contracts <- m_contracts %>%
  filter(Publish.Date >= start)
monthly_value <- sum(m_contracts$Value)
monthly_contracts <- dim(m_contracts)[1]

```
   
Total value of contracts  = $`r total_value`  
Total number of contracts = `r total_contracts`   
Monthly value of contracts = $`r monthly_value`  
Monthly number of contracts = `r monthly_contracts`  

  
## Sellers  
```{r,echo=FALSE}
m_sellers <- sellers %>%
  filter(joined <= end)
total_sellers <- dim(m_sellers)[1]

m_sellers <- m_sellers %>%
  filter(joined >= start)
monthly_sellers <- dim(m_sellers)[1]
```
  
Total sellers = `r total_sellers`  
Sellers joining in month = `r monthly_sellers`  
  
## Published briefs  
```{r,echo=FALSE}
m_briefs <- briefs %>%
  filter(published <= end)
total_briefs <- dim(m_briefs)[1]

open_to_selected <- m_briefs %>%
  group_by(openTo) %>%
  summarise(count = length(openTo))
open_to_selected$perc <- open_to_selected$count / sum(open_to_selected$count) * 100

type_counts <- m_briefs %>%
  group_by(type) %>%
  summarise(count = length(type))
type_counts$perc <- type_counts$count / sum(type_counts$count) * 100

m_briefs <- m_briefs %>%
  filter(published >= start)
monthly_briefs <- dim(m_briefs)[1]
```
  
Total briefs = `r total_briefs`  
Monthly briefs = `r monthly_briefs`  
  
Open to:  
`r kable(open_to_selected)`  
  
Brief type:  
`r kable(type_counts)`  
  
## Types of briefs  
Total briefs in each category since launch

```{r,echo=FALSE}
aoe_counts <- briefs %>%
  filter(published <= end, type=="Specialist") %>%
  group_by(areaOfExpertise) %>%
  summarise(count = length(areaOfExpertise)) %>%
  arrange(desc(count))
aoe_counts$areaOfExpertise <- as.character(aoe_counts$areaOfExpertise)
# sum of counts excluding the top 4 
sum_others <- sum(aoe_counts[5:dim(aoe_counts)[1],]$count)
# number of Outcomes
outcome_count <- briefs %>%
    filter(published <= end, type=="Outcome") %>%
    dim
aoe_counts <- rbind(aoe_counts,
                        c("Total others (all but top 4 specialist)",sum_others))
aoe_counts <- rbind(aoe_counts,
                    c("Total Outcomes",outcome_count))
```
`r kable(aoe_counts)`  
  
## Areas of expertise offered  

  
```{r,echo=FALSE}
m_s   <- sellers %>%
  filter(joined < end)
aoes <- sapply(domains,function(x,s) {
        sum(grepl(x,s$assessed_aoe))
       },s = m_s)
aoes_un <- sapply(domains,function(x,s) {
        sum(grepl(x,s$unassessed_aoe))
       },s = m_s)
aoes <- aoes + aoes_un
aoe_df <- data.frame(count=aoes)
```

`r kable(aoe_df)`  

## Contracts by seller  

```{r,echo=FALSE}
m_c <- contracts %>%
  filter(Publish.Date <= end, Publish.Date >= fy_start)
fy_contracts <- m_c %>%
  group_by(Supplier.ABN) %>%
  summarise(value = sum(Value),count=length(Supplier.ABN))
fy_contracts <- merge(fy_contracts,sellers[,c("abn","name","sme_by_employees")],
                      by.x="Supplier.ABN",by.y="abn",all.x=TRUE) %>%
  arrange(desc(count),desc(value))
fy_others <- fy_contracts[11:dim(fy_contracts)[1],] %>%
  group_by(sme_by_employees) %>%
  summarise(value = sum(value),count = sum(count))

  
```
### Current financial year  
#### Top ten sellers
`r kable(fy_contracts[1:10,])`  
#### Count of 'others' - excludes the top 10 sellers  
`r kable(fy_others)`  





