---
title: "Marketplace Monthly Performance"
output:
  html_document:
    theme: journal
    toc: true
    toc_float: true
    df_print: paged
  pdf_document: default
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!exists("buyers")) {
  source("../R/utilities.r")
  source("../R/plotly_functions.r")
  source("../R/assorted_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  library(httr)
  library(stringr)
  library(kableExtra)
  library(lubridate)
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),rel_path_data(),substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
}

month <- "month" # change to "week" to get a weekly report  

current_month   <- floor_date(Sys.Date(),unit=month) # filter out events after this date as month
reporting_month <- floor_date(current_month-1,unit=month) # month in focus
previous_month  <- floor_date(reporting_month-1,unit=month)
reporting_month_name  <- as.character(month(reporting_month,label=TRUE,abbr=FALSE))
previous_month_name   <- as.character(month(previous_month,label=TRUE,abbr=FALSE))


c_m    <- contracts %>% 
  filter(Publish.Date < current_month) %>%
  mutate(month = floor_date(Publish.Date,unit=month))
b_m    <- briefs %>% 
  mutate(publish_month = floor_date(published,unit=month),
         close_month   = floor_date(close,unit=month)) %>%
  filter(published < current_month)
```   

```{r,echo=FALSE}  
b <- b_m %>%
  group_by(publish_month) %>%
  summarise(count = n(),
            publishing_agencies = length(unique(email_domain)),
            all = sum(openTo == "All")/n()*100,
            some = sum(openTo == "Some")/n()*100,
            one  = sum(openTo == "One")/n()*100,
            specialist = sum(type == "Specialist")/n()*100,
            outcome    = sum(type == "Outcome")/n()*100,
            training   = sum(type == "Training")/n()*100,
            rfx        = sum(type == "RFX")/n()*100,
            comm_cce   = sum(entities == "CCE")/n()*100,
            local      = sum(entities == "Local")/n()*100,
            comm_ncce  = sum(entities == "NCCE")/n()*100,
            other      = sum(entities == "Other")/n()*100,
            state      = sum(entities == "State")/n()*100,
            all_total  = sum(openTo == "All"),
            some_total = sum(openTo == "Some"),
            one_total  = sum(openTo == "One"),
            specialist_total = sum(type == "Specialist"),
            outcome_total    = sum(type == "Outcome"),
            training_total   = sum(type == "Training"),
            rfx_total        = sum(type == "RFX"),
            comm_cce_total   = sum(entities == "CCE"),
            local_total      = sum(entities == "Local"),
            comm_ncce_total  = sum(entities == "NCCE"),
            other_total      = sum(entities == "Other"),
            state_total      = sum(entities == "State")
            )
b_rep  <- b[which(b$publish_month==reporting_month),]
b_prev <- b[which(b$publish_month==previous_month),]
```  

# Briefs   
<br />
**Summary for `r reporting_month_name`**    

*  `r b_rep$count[1]` briefs were published (`r b_prev$count[1]` published in `r previous_month_name`)    
*  `r b_rep$publishing_agencies[1]` agencies published briefs (`r b_prev$publishing_agencies[1]` published in `r previous_month_name`)  


## Number of briefs published  
<br />  
```{r,echo=FALSE}
b_lo <- b %>% 
  mutate(month = seq_along(publish_month)) %>%
  loess(count ~ month,data=.,span=0.4)
b$smoothed <- b_lo$fitted

p <- plot_ly(data=b,x=~publish_month,y=~count,type="scatter",mode="lines+markers",name="briefs") %>%
  add_trace(y=~smoothed,name="briefs trend") %>%
  layout(title = "Count of briefs published each month",
         xaxis = list(title="month published"),
         yaxis = list(title="briefs published",range=c(0,max(b$count))))
#p
```
`r p`  
  
  
## Split of briefs by who open to  
<br />  
```{r,echo=FALSE}  
#by percentage
p1 <- plot_ly(data=b,x=~publish_month,y=~all,name="all",type="scatter",mode="lines+markers") %>%
  add_trace(y=~some,name="some") %>%
  add_trace(y=~one,name="one") %>%
    layout(title = "Percentage of briefs by sellers open to",
         xaxis = list(title=""),
         yaxis = list(title="percentage",range=c(0,100)),
         legend=list(orientation="h"))
# by value
p2 <- plot_ly(data=b,x=~publish_month,y=~all_total,name="all",type="scatter",mode="lines+markers") %>%
  add_trace(y=~some_total,name="some") %>%
  add_trace(y=~one_total,name="one") %>%
  add_trace(y=~count,name="total") %>%
    layout(title = "Number of briefs by sellers open to",
         xaxis = list(title=""),
         yaxis = list(title="number of briefs"),
         legend=list(orientation="h"))

```
  
`r p1`  
<br />  
`r p2`  

## Split of briefs by type  
<br />  

```{r,echo=FALSE}  
p3 <- plot_ly(data=b,x=~publish_month,y=~outcome,name="outcomes",type="scatter",mode="lines+markers") %>%
  add_trace(y=~specialist,name="specialists") %>%
  add_trace(y=~training,name="training") %>%
  add_trace(y=~rfx,name="RFX") %>%
    layout(title = "Percentage of briefs by type",
         xaxis = list(title=""),
         yaxis = list(title="percentage",range=c(0,100)),
         legend=list(orientation="h"))

p4 <- plot_ly(data=b,x=~publish_month,y=~outcome_total,name="outcomes",type="scatter",mode="lines+markers") %>%
  add_trace(y=~specialist_total,name="specialists") %>%
  add_trace(y=~training_total,name="training") %>%
  add_trace(y=~rfx_total,name="RFX") %>%
  add_trace(y=~count,name="total") %>%
    layout(title = "Number of briefs by type",
         xaxis = list(title=""),
         yaxis = list(title="number of briefs"),
         legend=list(orientation="h"))

```

`r p3`  
<br />  
`r p4`   

# Buyers  

## Agencies publishing  
<br />  
```{r,echo=F}
p5 <- plot_ly(data=b,x=~publish_month,y=~publishing_agencies,type="scatter",mode="lines+markers") %>%
  layout(title = "Count of agencies publishing in each month",
         xaxis = list(title=""),
         yaxis = list(title="agencies publishing briefs"))  
```   
`r p5`  

## Split of briefs by publishing entity type  
<br />  

```{r,echo=FALSE}  
p6 <- plot_ly(data=b,x=~publish_month,y=~comm_ncce,name="Comm NCCE",type="scatter",mode="lines+markers") %>%
  add_trace(y=~comm_cce,name="Comm CCE") %>%
  add_trace(y=~state,name="State") %>%
  add_trace(y=~local,name="Local") %>%
  add_trace(y=~other,name="Other") %>%
    layout(title = "Percentage of briefs by publishing entity type",
         xaxis = list(title=""),
         yaxis = list(title="percentage",range=c(0,100)),
         legend=list(orientation="h"))

p7 <- plot_ly(data=b,x=~publish_month,y=~comm_ncce_total,name="Comm NCCE",type="scatter",mode="lines+markers") %>%
  add_trace(y=~comm_cce_total,name="Comm CCE") %>%
  add_trace(y=~state_total,name="State") %>%
  add_trace(y=~local_total,name="Local") %>%
  add_trace(y=~other_total,name="Other") %>%
  add_trace(y=~count,name="Total") %>%
    layout(title = "Number of briefs by publishing entity type",
         xaxis = list(title=""),
         yaxis = list(title="number of briefs",range=c(0,100)),
         legend=list(orientation="h"))

```  
`r p6`  
<br />  
`r p7`  

---  


# Contracts  
## Briefs vs Contracts  
```{r,echo=FALSE}
b_ncce <- b_m %>%
  filter(entities == "NCCE") %>%
  group_by(publish_month) %>%
  summarise(count = n())

b_lo <- b_ncce %>% 
  mutate(month = seq_along(publish_month)) %>%
  loess(count ~ month,data=.,span=0.4)
b_ncce$smoothed <- b_lo$fitted

c_summary <- c_m %>% 
  group_by(month) %>%
  summarise(count=n()) %>%
  ungroup() %>%
  mutate(month_seq = seq_along(month))

c_smooth <- c_summary %>%  
  loess(count ~ month_seq,data=.,span=0.4)

c_summary$smoothed <- c_smooth$fitted

p <- plot_ly(data=b_ncce,x=~publish_month,y=~count,
             type="scatter",mode="lines+markers",name="briefs",
             opacity=0.4) %>%
  add_trace(y=~smoothed,name="briefs trend",opacity=1.0) %>%
  add_trace(x=c_summary$month,y=c_summary$count,name="contracts",opacity=0.4) %>%
  add_trace(x=c_summary$month,y=c_summary$smoothed,name="contracts trend",opacity=1.0) %>%
  layout(title = "Count of briefs and contracts published each month",
         xaxis = list(title="month published"),
         yaxis = list(title="briefs published",range=c(0,max(c_summary$count))))

```  
`r p`  

Only briefs published by non corporate commonwealth entities are shown, as these are the only agencies publishing contracts through Austender.  
  
  

## Percentage of contracts to SMEs by month   
<br />  
```{r,echo=FALSE}
c_cum <- c_m %>% 
  arrange(Publish.Date) %>%
  mutate(month = floor_date(Publish.Date,unit="month")) %>%
  group_by(month,sme_by_employees) %>%
  summarise(
    count = n(),
    value = sum(Value)
  ) %>%
  mutate(seller_type = case_when(
    sme_by_employees == TRUE ~ "SME",
    TRUE ~ "Enterprise"
  )) %>%
  ungroup() %>%
  select(-sme_by_employees)

by_count <- c_cum %>%
  select(-value) %>%
  spread(seller_type,count,fill=0) %>% 
  mutate(net_SME = cumsum(SME),
         net_Ent = cumsum(Enterprise),
         cumulative_SME_perc = net_SME/(net_SME+net_Ent)*100,
         monthly_SME_perc = SME /(SME+Enterprise)*100)

p8 <- plot_ly(data=by_count,x=~month,y=~cumulative_SME_perc,type="scatter",mode="lines+markers",name="Cumulative") %>%
  add_trace(y=~monthly_SME_perc,name="Monthly") %>%
  layout(title="Percentage of contracts awarded to SMEs",
         yaxis = list(title="Percentage of contracts awarded to SME"))

by_value <- c_cum %>%
  select(-count) %>%
  spread(seller_type,value,fill=0) %>% 
  mutate(net_SME = cumsum(SME),
         net_Ent = cumsum(Enterprise),
         cumulative_SME_perc = net_SME/(net_SME+net_Ent)*100,
         monthly_SME_perc = SME /(SME+Enterprise)*100)

p9 <- plot_ly(data=by_value,x=~month,y=~cumulative_SME_perc,type="scatter",mode="lines+markers",name="Cumulative") %>%
  add_trace(y=~monthly_SME_perc,name="Monthly") %>%
  layout(title="Percentage of contract value awarded to SMEs",
         yaxis = list(title="Percentage of contract value awarded to SME"))
```  
`r p8`  
<br />  
`r p9`  

## 
## Top buyers publishing  
## Top buyers by contracts published  
## New agencies this month  

