---
title: "Monthly Insights Reports Data Summary"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## set manual_date = TRUE to specify the dates manually. Otherwise the
# dates will default to the last full month based on the Sys.Date
manual_date = FALSE
if (manual_date) {
  start <- as.Date("2018-09-01")
  end   <- as.Date("2018-09-31")
} else {
  end   <- floor_date(Sys.Date(),unit="month") - 1
  start <- floor_date(end,unit="month")
}

fy_start <- as.Date("2018-07-01")
ninety_days <- seq(end-89,end,by="day")
#reference_file <- "..\\data\\2017-09-19.Rdata"

# if rendered in a new R session, load the session values
if (!exists("buyers")) {
  source("utilities.r")
  source("plotly_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  ## load the files - ensure these files are already produced
  #load(paste0(getwd(),"/../data/",substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
  load(paste0(getwd(),"\\..",rel_path_data,substr(as.POSIXct(Sys.time()),1,10),".Rdata"))

  #load(reference_file)
}


```  
Date range for report:  
Starts `r start`  
Ends `r end`  
Current financial year: `r fy_start`  
  
# Who is buying?

```{r,echo=FALSE}
agencies <- buyers %>%
    group_by(entities) %>% 
    summarise(count=length(unique(email_domain)))
cats <- c("Commonwealth - CCE",
            "Commonwealth - NCCE",
            "Local","State / Territory","Other")
agencies$entities <- translate(
                        qw("CCE NCCE Local State Other"),
                        cats,
                        agencies$entities)
agencies <- agencies %>% arrange(entities)        
#agencies_counts <- buyers %>%
#                group_by(email_domain,category) %>% 
#                summarise(count=length(email_domain)) %>%
#                group_by(category) %>%
#                summarise(count = length(category))
#agencies_counts$percent <- round(agencies_counts$count/sum(agencies_counts$count) * 100,digits=1)
#agencies_counts <- rbind(agencies_counts,c("Total agencies",sum(agencies_counts$count),100.0))

plot_agency_count_by_category_2(agencies)

agencies <- rbind(agencies,
                  c("Total number of registered entities",sum(agencies$count)))

```  

`r kables(agencies)`   
  
  
  
## Top buyers in month  

```{r,echo=FALSE}
buyer_counts <- briefs %>%
  filter(published >= start,published <= end) %>%
  group_by(agencyName) %>%
  summarise(count=length(agencyName)) %>%
  arrange(desc(count))
```

`r kables(buyer_counts)`     

**Total buyers publishing in month `r dim(buyer_counts)[1]`**  


  
# How are we encouraging competition?  
## Open to      
```{r,echo=FALSE}
x <- briefs %>% group_by(openTo) %>% summarise(count = length(openTo))
plot_ly(x,labels=~openTo, values=~count,
          textposition = 'outside',
          textinfo = 'label+percent+value',
          marker=list(colors = brewer.pal(5,"Dark2"),
                      line = list(color = '#FFFFFF', width = 1))) %>%
    add_pie(hole=0.4) %>%
    layout(title = "Open to",
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```  
  
## Types of briefs published    
```{r,echo=FALSE}
x <- briefs %>% 
  mutate(type = as.character(type)) %>%
  #mutate(type = case_when(
  #                type == "Outcome" & openTo == "All" ~ "ATM",
  #                type == "Outcome"                   ~ "RFX",
  #                TRUE                                ~ type
  #              )
  #       ) %>%
  group_by(type) %>% 
  summarise(count = length(type))
plot_ly(x,labels=~type, values=~count,
          textposition = 'outside',
          textinfo = 'label+percent+value',
          marker=list(colors = brewer.pal(5,"Dark2"),
                      line = list(color = '#FFFFFF', width = 1))) %>%
    add_pie(hole=0.4) %>%
    layout(title = "Types of brief",
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```  

## % dollar value awarded to SMEs tile    


```{r,echo=FALSE}
# sme figures
#m_contracts <- left_join(contracts,sellers,by=c("Supplier.ABN" = "abn"))
m_contracts <- contracts

sme_emps  <- m_contracts %>% 
  filter(Publish.Date <= end) %>%
  group_by(sme_by_employees) %>% 
  summarise(value = sum(Value), count = length(sme_by_employees))
sme_emps$by_value <- sme_emps$value / sum(sme_emps$value) * 100
sme_emps$by_count <- sme_emps$count / sum(sme_emps$count) * 100
```
`r round(sme_emps[sme_emps$sme_by_employees==T,"by_value"],digits=0)`% of the dollar value contracted through the Marketplace...  


`r kables(sme_emps)`  
  
## Contract values tile  

```{r,echo=FALSE}
# all contracts awarded to the end of the reporting month
m_contracts <- contracts %>%
  filter(Publish.Date <= end)
total_value <- round(sum(m_contracts$Value)/1000000,digits=2)
total_contracts <- dim(m_contracts)[1]

# filter out contracts awarded prior to the reporting month
m_contracts <- m_contracts %>%
  filter(Publish.Date >= start)
monthly_value <- round(sum(m_contracts$Value)/1000000,digits=2)
monthly_contracts <- dim(m_contracts)[1]

m_by_sme <- m_contracts %>%
  group_by(sme_by_employees) %>%
  summarise(count = length(sme_by_employees), value = sum(Value))
m_by_sme$perc_by_count <- m_by_sme$count / (sum(m_by_sme$count)) * 100
m_by_sme$perc_by_value <- m_by_sme$value / (sum(m_by_sme$value)) * 100

```

Total value of contracts  = $`r total_value`M  
  
<br />  
To determine the contract value for "this month", deduct the total contract value in last month's report from this month's total (`r total_value`).   
Value of *new* contracts this month = $`r monthly_value`M (but this doesn't include variations to old contracts) 

### Spark line of contract value in the past 90 days  

```{r,echo=FALSE}
# show a spark line of contract value over the past 90 days
m_contracts <- contracts %>%
  filter(Publish.Date > end - 90) %>%
  arrange(Publish.Date) %>%
  mutate(cumulative = cumsum(Value))
vals <- sapply(ninety_days,function(x,c) {
                c <- c %>% filter(c$Publish.Date <= x)
                if (dim(c)[1] == 0) {
                  return(0)
                }
                return(max(c$cumulative))
              },c=m_contracts)
plot_ly(x=~ninety_days,y=~vals,mode="lines",type="scatter") %>%
  layout(title="",
         xaxis=list(visible=FALSE),
         yaxis=list(visible=FALSE))
```  

## Sellers tile    
```{r,echo=FALSE}
m_sellers <- sellers %>%
  filter(joined <= end)
total_sellers <- dim(m_sellers)[1]

m_sellers <- m_sellers %>%
  filter(joined >= start)
monthly_sellers <- dim(m_sellers)[1]
```
  
Total sellers = `r total_sellers`  
Sellers joining in month = `r monthly_sellers`  
  
### Spark line of registered sellers in the past 90 days  

```{r,echo=FALSE}
# show a spark line of contract value over the past 90 days
s <- sellers %>%
  filter(joined > end - 90, joined <= end) %>%
  arrange(joined)
s$count <- 1:dim(s)[1]
vals <- sapply(ninety_days,function(x,s) {
                s <- s %>% filter(s$joined <= x)
                if (dim(s)[1] == 0) {
                  return(0)
                }
                return(max(s$count))
              },s=s)
plot_ly(x=~ninety_days,y=~vals,mode="lines",type="scatter") %>%
  layout(title="",
         xaxis=list(visible=FALSE),
         yaxis=list(visible=FALSE))
```    
  
## Opportunities tile    
```{r,echo=FALSE}
ops <- briefs %>%
  filter(published <= end)
total_ops <- dim(ops)[1]

ops <- ops %>%
  filter(published >= start)
monthly_ops <- dim(ops)[1]
```
  
Total opportunities = `r total_ops`  
Opportunities in the month = `r monthly_ops`  
  
### Spark line of opportunities in the past 90 days  

```{r,echo=FALSE}
b <- briefs %>%
  filter(published > end - 90, published <= end) %>%
  arrange(published)
b$count <- 1:dim(b)[1]
vals <- sapply(ninety_days,function(x,b) {
                b <- b %>% filter(b$published <= x)
                if (dim(b)[1] == 0) {
                  return(0)
                }
                return(max(b$count))
              },b=b)
plot_ly(x=~ninety_days,y=~vals,mode="lines",type="scatter") %>%
  layout(title="",
         xaxis=list(visible=FALSE),
         yaxis=list(visible=FALSE))
``` 

# How do we break down procurement  

```{r,echo=FALSE}
by_phase <- briefs %>% 
  filter(!is.na(phase)) %>%
  group_by(phase) %>%
  summarise(count = length(phase))
by_phase$percentage <- round(by_phase$count/sum(by_phase$count) * 100)
by_phase <- rbind(by_phase,c("Total",sum(by_phase$count),100))
NAs      <- sum(is.na(briefs$phase))
```  
  
`r kables(by_phase)`  
Number of briefs with phase not set: `r NAs` - these are not counted above  

  
## Trend for day rates  
```{r,echo=FALSE}
br <- briefResponses %>% filter(applicationDate <= end, areaOfExpertise != "Change  Training and Transformation")
plot_violins_of_day_rates(br)
```  


## Types of briefs  
Total briefs in each category since launch

```{r,echo=FALSE}
aoe_counts <- briefs %>%
  filter(published <= end, type=="Specialist") %>%
  group_by(areaOfExpertise) %>%
  summarise(count = length(areaOfExpertise)) %>%
  arrange(desc(count))
aoe_counts$areaOfExpertise <- as.character(aoe_counts$areaOfExpertise)
# sum of counts excluding the top 4 
sum_others <- sum(aoe_counts[6:dim(aoe_counts)[1],]$count)
# number of Outcomes
outcome_count <- briefs %>%
    filter(published <= end, type=="Outcome") %>%
    dim
aoe_counts <- rbind(aoe_counts,
                        c("Total others (all but top 5 specialist)",sum_others))
aoe_counts <- rbind(aoe_counts,
                    c("Total Outcomes",outcome_count))
aoe_top5   <- aoe_counts[c(1:5,12,13),]
plot_ly(aoe_top5,
        labels=~areaOfExpertise,
        values=~count,
        type="pie",
        textinfo = 'label+percent+value',
        textposition = 'outside',
        sort=FALSE,
        direction="clockwise",
        marker=list(
          colors = brewer.pal(7,"Blues")))
```
`r kables(aoe_counts)`  

# How active is the Marketplace?  

## Areas of expertise offered  

  
```{r,echo=FALSE}
m_s   <- sellers %>%
  filter(joined < end)
aoes <- sapply(domains,function(x,s) {
        sum(grepl(x,s$assessed_aoe))
       },s = m_s)
aoes_un <- sapply(domains,function(x,s) {
        sum(grepl(x,s$unassessed_aoe))
       },s = m_s)
aoes <- aoes + aoes_un
aoe_df <- data.frame(count=aoes)
```

`r kables(aoe_df)`  
## Brief responses  
```{r,echo=FALSE}
#b_filt <- briefs %>% 
#  filter(openTo=="All",status=="closed")
#br <- briefResponses %>% filter(!is.na(supplierId),applicationDate <= end)
#b_filt$responses <- sapply(b_filt$id,function(x,br) {
#                                       sum(br$id==x)
#                                     },br=br)
#dens <- density(b_filt$responses,from=0)
#plot_ly(x=dens$x,y=dens$y,type="scatter",mode="lines") %>%
#    layout(title="Distribution of responses",
#         yaxis=list(visible=FALSE))

# now by specialists and outcomes, and only looking at the past 12 months
specialists <- briefResponses %>% 
  filter(openTo=="All", type=="Specialist") %>% 
  filter(published > Sys.Date() - 365) %>%
  group_by(id) %>% 
  summarise(responses = sum(attachments)) %>%
  filter(!is.na(responses))
outcomes <- briefResponses %>% 
  filter(openTo=="All", type=="Outcome") %>% 
  filter(published > Sys.Date() - 365) %>%
  group_by(id) %>% 
  summarise(responses = n()) %>%
  filter(!is.na(responses))
dens_s <- density(specialists$responses,from=0,to=120)
dens_s$y <- dens_s$y / (max(dens_s$y)) * 45
dens_o <- density(outcomes$responses,from=0,adjust=1.5)
dens_o$y <- dens_o$y / (max(dens_o$y)) * 45
plot_ly(alpha = 0.6) %>%
  add_histogram(x=specialists$responses,type="specialists",name="histogram") %>%
  add_histogram(x=outcomes$responses,type="outcomes",name="histogram") %>%
  add_lines(x=dens_s$x,y=dens_s$y,name="Specialists") %>%
  add_lines(x=dens_o$x,y=dens_o$y,name="Outcomes") %>%
  layout(
    barmode="overlay",
    title="Seller Responses per Brief - reference graph",
    xaxis=list(title="Applications"),
    yaxis=list(title="Count"))

#

```
Summary figures:  

Type | Min. | 1st Quartile | Median | Mean | 3rd Quartile | Max.  
--|--|--|--|--|--|--  
Specialists | `r paste(paste(summary(specialists$responses),collapse = " | "),"  ")`  
Outcomes | `r paste(paste(summary(outcomes$responses),collapse = " | "),"  ")`  

(Includes all Open to All and closed briefs in the past year)  


# Who is winning the work?  

```{r,echo=FALSE}
fy_contracts <- contracts %>%
  filter(Publish.Date <= end, Publish.Date >= fy_start) %>%
  group_by(Supplier.ABN,name,sme_by_employees) %>%
  summarise(value = sum(Value),
            count = n()) %>%
  arrange(desc(count),desc(value))

#fy_contracts <- m_c %>%
#  group_by(Supplier.ABN) %>%
#  summarise(value = sum(Value),count=length(Supplier.ABN))
#fy_contracts <- merge(fy_contracts,sellers[,c("abn","name","sme_by_employees")],
#                      by.x="Supplier.ABN",by.y="abn",all.x=TRUE) %>%
#  arrange(desc(count),desc(value))
fy_others <- fy_contracts[11:dim(fy_contracts)[1],] %>%
  group_by(sme_by_employees) %>%
  summarise(value = sum(value),count = sum(count)) %>%
  mutate(name=c("Others-Enterprise","Others-SME"))
#fy_all <- bind_rows(fy_contracts[1:10,],fy_others)
#plot_ly(fy_all,
#        labels=~name,
#        rotation="35",
#        values=~count,
#        type="pie",
#        textinfo = 'label+value',
#        textposition = 'outside',
#        sort=FALSE,
#        direction="clockwise"
        #marker = list(
        #  colors= colours)
#        ) %>% layout(showlegend=F)
```
  
### Current financial year  
#### Top ten sellers
`r kables(fy_contracts[1:10,])`  
#### Count of 'others' - excludes the top 10 sellers  
`r kables(fy_others)`  
  
## Contracts to SMEs for this month tile   

`r round(m_by_sme[m_by_sme$sme_by_employees==T,"perc_by_count"])`% of contracts have been awarded by volume to SME sellers this month  

`r kables(m_by_sme)`  
    
## Contracts this month tile  
  
`r monthly_contracts` Briefs have been contracted this month  






