---
title: "Marketplace Monthly Performance"
output:
  html_document:
    theme: journal
    toc: true
    toc_float: true
    df_print: paged
  pdf_document: default
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!exists("buyers")) {
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  library(httr)
  library(stringr)
  library(kableExtra)
  library(lubridate)
  library(googlesheets)
  source("../R/utilities.r")
  source("../R/plotly_functions.r")
  source("../R/assorted_functions.r")
  source("../R/scrape_austender.r")
  source("../R/process_functions.r")
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),rel_path_data(),substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
  cns <- load_austender_cns() %>% post_process_cns()
}

month <- "month" # change to "week" to get a weekly report  

current_month   <- floor_date(Sys.Date() + 6,unit=month) # filter out events after this date as month.
reporting_month <- floor_date(current_month-1,unit=month) # month in focus
previous_month  <- floor_date(reporting_month-1,unit=month)
reporting_month_name  <- as.character(month(reporting_month,label=TRUE,abbr=FALSE))
previous_month_name   <- as.character(month(previous_month,label=TRUE,abbr=FALSE))


c_m    <- contracts %>% 
  filter(Publish.Date < current_month) %>%
  mutate(month = floor_date(Publish.Date,unit=month))
b_m    <- briefs %>% 
  mutate(publish_month = floor_date(published,unit=month),
         close_month   = floor_date(close,unit=month)) %>%
  filter(published < current_month) # %>% 
  #mutate(openTo = case_when(type == "ATM" ~ "All",
  #                          TRUE          ~ openTo)) # ATMs are all open to all
b_m[b_m$type == "ATM",]$openTo <- "All" # mutate doesn't like the factors, so back to base
```   

```{r,echo=FALSE}  
b <- b_m %>%
  group_by(publish_month) %>%
  summarise(count = n(),
            publishing_agencies = length(unique(email_domain)),
            all = sum(openTo == "All")/n()*100,
            some = sum(openTo == "Some")/n()*100,
            one  = sum(openTo == "One")/n()*100,
            specialist = sum(type == "Specialist")/n()*100,
            outcome    = sum(type == "Outcome")/n()*100,
            training   = sum(type == "Training")/n()*100,
            rfx        = sum(type == "RFX")/n()*100,
            atm        = sum(type == "ATM")/n()*100,
            comm_cce   = sum(entities == "CCE")/n()*100,
            local      = sum(entities == "Local")/n()*100,
            comm_ncce  = sum(entities == "NCCE")/n()*100,
            other      = sum(entities == "Other")/n()*100,
            state      = sum(entities == "State")/n()*100,
            all_total  = sum(openTo == "All"),
            some_total = sum(openTo == "Some"),
            one_total  = sum(openTo == "One"),
            specialist_total = sum(type == "Specialist"),
            outcome_total    = sum(type == "Outcome"),
            training_total   = sum(type == "Training"),
            rfx_total        = sum(type == "RFX"),
            atm_total        = sum(type == "ATM"),
            comm_cce_total   = sum(entities == "CCE"),
            local_total      = sum(entities == "Local"),
            comm_ncce_total  = sum(entities == "NCCE"),
            other_total      = sum(entities == "Other"),
            state_total      = sum(entities == "State")
            )
b_rep  <- b[which(b$publish_month==reporting_month),]
b_prev <- b[which(b$publish_month==previous_month),]
```  

# Briefs   
<br />
**Summary for `r reporting_month_name`**    

*  `r b_rep$count[1]` briefs were published (`r b_prev$count[1]` published in `r previous_month_name`)    
*  `r b_rep$publishing_agencies[1]` agencies published briefs (`r b_prev$publishing_agencies[1]` published in `r previous_month_name`)  


## Number of briefs published  
<br />  
```{r,echo=FALSE}
b_lo <- b %>% 
  mutate(month = seq_along(publish_month)) %>%
  loess(count ~ month,data=.,span=0.4)
b$smoothed <- b_lo$fitted

p <- plot_ly(data=b,x=~publish_month,y=~count,type="scatter",mode="lines+markers",name="briefs") %>%
  add_trace(y=~smoothed,name="briefs trend") %>%
  layout(title = "Count of briefs published each month",
         xaxis = list(title="month published"),
         yaxis = list(title="briefs published"),
         legend=list(orientation="h"))
#p
```
`r p`  
  
  
## Split of briefs by who open to  
<br />  
```{r,echo=FALSE}  
#by percentage
p1 <- plot_ly(data=b,x=~publish_month,y=~all,name="all",type="scatter",mode="lines+markers") %>%
  add_trace(y=~some,name="some") %>%
  add_trace(y=~one,name="one") %>%
    layout(title = "Percentage of briefs by sellers open to",
         xaxis = list(title=""),
         yaxis = list(title="percentage",range=c(0,100)),
         legend=list(orientation="h"))
# by value
p2 <- plot_ly(data=b,x=~publish_month,y=~all_total,name="all",type="scatter",mode="lines+markers") %>%
  add_trace(y=~some_total,name="some") %>%
  add_trace(y=~one_total,name="one") %>%
  add_trace(y=~count,name="total") %>%
    layout(title = "Number of briefs by sellers open to",
         xaxis = list(title=""),
         yaxis = list(title="number of briefs"),
         legend=list(orientation="h"))

```
  
`r p1`  
<br />  
`r p2`  

## Split of briefs by type  
<br />  

```{r,echo=FALSE}  
p3 <- plot_ly(data=b,x=~publish_month,y=~outcome,name="outcomes",type="scatter",mode="lines+markers") %>%
  add_trace(y=~specialist,name="specialists") %>%
  add_trace(y=~training,name="training") %>%
  add_trace(y=~rfx,name="RFX") %>%
  add_trace(y=~atm,name="ATM") %>%
    layout(title = "Percentage of briefs by type",
         xaxis = list(title=""),
         yaxis = list(title="percentage",range=c(0,100)),
         legend=list(orientation="h"))

p4 <- plot_ly(data=b,x=~publish_month,y=~outcome_total,name="outcomes",type="scatter",mode="lines+markers") %>%
  add_trace(y=~specialist_total,name="specialists") %>%
  add_trace(y=~training_total,name="training") %>%
  add_trace(y=~rfx_total,name="RFX") %>%
  add_trace(y=~atm_total,name="ATM") %>%
  add_trace(y=~count,name="total") %>%
    layout(title = "Number of briefs by type",
         xaxis = list(title=""),
         yaxis = list(title="number of briefs"),
         legend=list(orientation="h"))

```

`r p3`  
<br />  
`r p4`   

# Buyers  

## Total agencies publishing  
<br />  
```{r,echo=F}
p5 <- plot_ly(data=b,x=~publish_month,y=~publishing_agencies,type="scatter",mode="lines+markers") %>%
  layout(title = "Count of agencies publishing in each month",
         xaxis = list(title=""),
         yaxis = list(title="agencies publishing briefs"))  
```   
`r p5`    


## Agencies publishing by last published  
```{r, echo = FALSE}
b_bym <- b_m %>% 
  select(id, email_domain, published) %>% 
  mutate(month = floor_date(published, unit = "month")) %>% 
  arrange(email_domain, published) %>% 
  group_by(email_domain) %>% 
  mutate(last_published       = lag(month),
         months_since_last    = interval(last_published, month) %/% months(1))

last_published <- b_bym %>% 
  group_by(month) %>%
  summarise(unique       = length(unique(email_domain)),
            last_month   = sum(months_since_last == 1, na.rm = TRUE),
            two_or_three = sum(months_since_last %in% c(2,3), na.rm = TRUE),
            earlier      = sum(months_since_last > 3, na.rm = TRUE),
            first_time   = sum(is.na(months_since_last)))

not_seen_since <- b_bym %>% 
  group_by(email_domain) %>% 
  mutate(latest = max(month), 
         months_since_latest = interval(latest, current_month) %/% months(1) - 1) %>% 
  ungroup() %>% 
  filter(months_since_latest > 2) %>% 
  group_by(latest) %>% 
  summarise(count = - length(unique(email_domain)))
    
p10 <- plot_ly(data = last_published, x =~ month, y =~ last_month, name = "published last month",
             type = "bar") %>% 
  add_trace(y =~ two_or_three, name = "2 or 3 months since last") %>% 
  add_trace(y =~ earlier, name = "4+ months since last") %>% 
  add_trace(y =~ first_time, name = "first time publisher") %>% 
  layout(title = "Agencies publishing each month by when last published",
         yaxis = list(title = "agencies publishing"),
         barmode = "stack")
```
`r p10`  

## Split of briefs by publishing entity type  
<br />  

```{r,echo=FALSE}  
p6 <- plot_ly(data=b,x=~publish_month,y=~comm_ncce,name="Comm NCCE",type="scatter",mode="lines+markers") %>%
  add_trace(y=~comm_cce,name="Comm CCE") %>%
  add_trace(y=~state,name="State") %>%
  add_trace(y=~local,name="Local") %>%
  add_trace(y=~other,name="Other") %>%
    layout(title = "Percentage of briefs by publishing entity type",
         xaxis = list(title=""),
         yaxis = list(title="percentage",range=c(0,100)),
         legend=list(orientation="h"))

p7 <- plot_ly(data=b,x=~publish_month,y=~comm_ncce_total,name="Comm NCCE",type="scatter",mode="lines+markers") %>%
  add_trace(y=~comm_cce_total,name="Comm CCE") %>%
  add_trace(y=~state_total,name="State") %>%
  add_trace(y=~local_total,name="Local") %>%
  add_trace(y=~other_total,name="Other") %>%
  add_trace(y=~count,name="Total") %>%
    layout(title = "Number of briefs by publishing entity type",
         xaxis = list(title=""),
         yaxis = list(title="number of briefs"),
         legend=list(orientation="h"))

```  
`r p6`  
<br />  
`r p7`  

---  


# Contracts   
```{r, echo = FALSE, results = 'asis'}
if (max(cns$end_date) < (current_month - 1)) {
  cat(paste("<p style='color:red'>__Contract data doesn't extend to the end of the month. Only contracts reported through to", max(cns$end_date), "are included below.__  </p>"))
}
```  
  
## Cumulative contract counts  
<br />  
```{r, echo = FALSE}
unique_smes_contracted <- c_m %>% 
  filter(sme_by_employees) %>%  
  group_by(Supplier.ABN) %>% 
  summarise(month = min(month)) %>% 
  ungroup() %>% 
  group_by(month) %>% 
  summarise(count_unique = n()) %>% 
  mutate(cumulative_unique_smes = cumsum(count_unique))

c_cum <- c_m %>% 
  group_by(month, sme_by_employees) %>% 
  summarise(count        = n(),
            unique_count = length(unique(Supplier.ABN))) %>%
  pivot_wider(id_cols    = month, 
              names_from = sme_by_employees, 
              values_from = c("count", "unique_count"),
              values_fill  = list(count = 0, unique_count = 0)) %>% 
  rename(count_sme  = count_TRUE,
         count_ent  = count_FALSE,
         unique_sme = unique_count_TRUE,
         unique_ent = unique_count_FALSE) %>% 
  ungroup() %>% 
  mutate(count_total  = count_sme + count_ent,
         unique_total = unique_sme + unique_ent,
         cumulative_sme = cumsum(count_sme),
         cumulative_ent = cumsum(count_ent),
         cumulative_total = cumsum(count_total)) %>% 
  left_join(select(unique_smes_contracted, month, cumulative_unique_smes), 
            by = "month") %>% 
  replace_na(list(cumulative_unique_smes = 0))

p <- plot_ly(data = c_cum, x =~ month, y =~ cumulative_total, 
             type = "scatter", mode = "lines+markers", name = "total contracts") %>% 
  add_trace(y =~ cumulative_sme, name = "sme contracts") %>% 
  add_trace(y =~ cumulative_unique_smes, name = "unique sme's contracted") %>% 
  layout(title = "Total contracts reported",
         xaxis = list(title = ""),
         yaxis = list(title = "number of contracts"),
         legend = list(orientation="h"))
```
`r p`  
`r max(c_cum$cumulative_total)` contracts have been reported.  
`r max(c_cum$cumulative_unique_smes)` unique SME's have been awarded `r max(c_cum$cumulative_sme)` of these contracts.   


## Cumulative contract values  

```{r, echo = FALSE, include = FALSE}
# NCCE ratio
ratio <- b_m %>% 
  arrange(published) %>% 
  group_by(publish_month) %>% 
  summarise(ncces = sum(entities == "NCCE"),
            briefs = n()) %>% 
  ungroup() %>% 
  mutate(ratio = cumsum(ncces) / cumsum(briefs))

cns_ <- cns %>%
  filter(SON.ID %in% dmp_sons) %>% 
  process_contract_records() 

cns_summ <- cns_ %>% 
  mutate(month     = as.Date(floor_date(Record.Published, unit = month)),
         amendment = (Amendment != 0)) %>% 
  filter(month < current_month) %>% 
  group_by(month, entities, amendment) %>% 
  summarise(count       = n(),
            value       = sum(value_amendment)) %>% 
  ungroup() %>% 
  pivot_wider(id_cols     = month,
              names_from  = c(entities, amendment), 
              values_from = c(count, value))  %>% 
  left_join(
    select(ratio, publish_month, ratio),
    by = c("month" = "publish_month")
  ) %>%   
  replace(is.na(.), 0) %>% 
  mutate(ratio = if_else(ratio == 0, 1, ratio)) %>% 
  mutate(cumulative_CCE      = cumsum(value_CCE_TRUE  + value_CCE_FALSE),
         cumulative_NCCE     = cumsum(value_NCCE_TRUE + value_NCCE_FALSE),
         monthly_amendment   = value_CCE_TRUE    + value_NCCE_TRUE,
         monthly_contract    = value_CCE_FALSE   + value_NCCE_FALSE,
         monthly_total       = monthly_amendment + monthly_contract,
         monthly_amendment_c = count_CCE_TRUE    + count_NCCE_TRUE,
         monthly_contract_c  = count_CCE_FALSE   + count_NCCE_FALSE,
         monthly_total_c     = monthly_amendment_c + monthly_contract_c,
         cumulative_total    = cumulative_CCE + cumulative_NCCE,
         adjusted_total      = cumulative_NCCE / ratio)

p <- plot_ly(data = cns_summ, x =~ month, y =~ adjusted_total, 
             mode = "lines+markers", type = "scatter", name = "adjusted total") %>% 
  add_trace(y =~ cumulative_total, name = "reported total") %>% 
  add_trace(y =~ cumulative_NCCE, name  = "NCCE contracts") %>% 
  add_trace(y =~ cumulative_CCE, name   = "CCE contracts")  %>% 
  layout(title = "Total value of reported contracts",
         xaxis = list(title = ""),
         yaxis = list(title = "value of contracts (AUD)"),
         legend = list(orientation="h"))

```
`r p`  
  
  
## Contract values by month  
<br />  
```{r, echo = FALSE}

p <- plot_ly(data=cns_summ,x =~ month, y =~ monthly_total,
        name="total",
        type="scatter",
        mode="lines+markers") %>%
  add_trace(y =~ monthly_contract,     name="new contracts") %>%
  add_trace(y =~ monthly_amendment,    name="variations") %>%
  layout(title  = "Contract values reported by month",
         yaxis  = list(title="Value ($M)"),
         xaxis  = list(title = ""),
         legend = list(orientation="h"))


```  

`r p`  
  
## Contract volumes  
<br />  

```{r, echo = FALSE}

p <- plot_ly(data = cns_summ, x =~ month, y =~ monthly_contract_c,
             type = "scatter",
             mode = "lines+markers",
             name = "total") %>%
  #add_trace(y =~ monthly_contract_c, name = "new contracts") %>% 
  #add_trace(y =~ monthly_, name = "amendments") %>% 
  layout(title="Contracts reported by month",
         yaxis  = list(title="count"),
         xaxis  = list(title = ""),
         legend = list(orientation="h"))


```  
`r p`  
  
  

## Briefs vs Contracts  
```{r,echo=FALSE}
b_ncce <- b_m %>%
  filter(entities == "NCCE") %>%
  group_by(publish_month) %>%
  summarise(count = n())

b_lo <- b_ncce %>% 
  mutate(month = seq_along(publish_month)) %>%
  loess(count ~ month,data=.,span=0.4)
b_ncce$smoothed <- b_lo$fitted

c_summary <- c_m %>% 
  group_by(month) %>%
  summarise(count=n()) %>%
  ungroup() %>%
  mutate(month_seq = seq_along(month))

c_smooth <- c_summary %>%  
  loess(count ~ month_seq,data=.,span=0.4)

c_summary$smoothed <- c_smooth$fitted

p <- plot_ly(data=b_ncce,x=~publish_month,y=~count,
             type="scatter",mode="lines+markers",name="briefs",
             opacity=0.4) %>%
  add_trace(y=~smoothed,name="briefs trend",opacity=1.0) %>%
  add_trace(x=c_summary$month,y=c_summary$count,name="contracts",opacity=0.4) %>%
  add_trace(x=c_summary$month,y=c_summary$smoothed,name="contracts trend",opacity=1.0) %>%
  layout(title = "Count of briefs and contracts published each month",
         xaxis = list(title="month published"),
         yaxis = list(title="count"))

```  
`r p`  

Only briefs published by non corporate commonwealth entities (NCCE) are shown, as these are the only agencies publishing contracts through Austender.  
  
## Marketplace panel vs other panels   
  
```{r,echo=FALSE}
panels_by_month <- filter_cns_to_latest_amendment(cns) %>%
  filter(str_detect(SON.ID,"SON"), month < current_month, month > ad("2016-07-31")) %>% 
  select(CN.ID,SON.ID,Publish.Date,month,Value) %>%
  mutate(dmp_panel = SON.ID %in% dmp_sons) %>%
  group_by(month,dmp_panel) %>%
  summarise(count = n(),
            value = sum(Value))

panels_by_count <- panels_by_month %>%
  select(-value) %>%
  spread(key = dmp_panel, value = count, fill = 0) %>%
  rename(dmp = `TRUE`, all_sons = `FALSE`) %>%
  ungroup()

panels_by_count$dmp_smoothed <- loess(dmp ~ seq_along(month),
                                      data = panels_by_count,
                                      span = 0.4)$fitted
panels_by_count$all_sons_smoothed <- loess(all_sons ~ seq_along(month),
                                           data = panels_by_count,
                                           span = 0.4)$fitted

ay <- list(
  #tickfont = list(color = "red"),
  overlaying = "y",
  side = "right",
  title = "All panel contracts including DMP"
)
colours = c("light blue", "blue", "light red", "red")
plot_ly() %>%
  add_lines(x =~ month, y =~ dmp, data = panels_by_count, name = "DMP",opacity = 0.4) %>%
  add_lines(x = panels_by_count$month, y = panels_by_count$dmp_smoothed, 
            name = "DMP trend", opacity = 1.0) %>%
  add_lines(x =~ month, y =~ all_sons, name = "All SONs", 
             opacity = 0.4) %>% #yaxis = "y2",
  add_lines(x =~ month, y =~ all_sons_smoothed, name = "All trend", 
            opacity = 1.0) %>% #yaxis = "y2", 
  layout(
    title = "DMP contracts versus all other panels", yaxis2 = ay,
    xaxis = list(title=""),
    yaxis = list(title="count of dmp contracts only", type = "log"))

```  


## Percentage of contracts to SMEs by month   
<br />  
```{r,echo=FALSE}
c_cum <- c_m %>% 
  arrange(Publish.Date) %>%
  mutate(month = floor_date(Publish.Date,unit="month")) %>%
  group_by(month,sme_by_employees) %>%
  summarise(
    count = n(),
    value = sum(Value)
  ) %>%
  mutate(seller_type = case_when(
    sme_by_employees == TRUE ~ "SME",
    sme_by_employees == FALSE ~ "Enterprise",
    TRUE ~ "unknown"
  )) %>%
  ungroup() %>%
  select(-sme_by_employees)

by_count <- c_cum %>%
  select(-value) %>%
  spread(seller_type,count,fill=0) %>% 
  mutate(net_SME = cumsum(SME),
         net_Ent = cumsum(Enterprise),
         cumulative_SME_perc = net_SME/(net_SME+net_Ent)*100,
         monthly_SME_perc = SME /(SME+Enterprise)*100)

p8 <- plot_ly(data=by_count,x=~month,y=~cumulative_SME_perc,type="scatter",mode="lines+markers",name="Cumulative") %>%
  add_trace(y=~monthly_SME_perc,name="Monthly") %>%
  layout(title="Percentage of contracts awarded to SMEs",
         yaxis = list(title="Percentage of contracts awarded to SME"))

by_value <- c_cum %>%
  select(-count) %>%
  spread(seller_type,value,fill=0) %>% 
  mutate(net_SME = cumsum(SME),
         net_Ent = cumsum(Enterprise),
         cumulative_SME_perc = net_SME/(net_SME+net_Ent)*100,
         monthly_SME_perc = SME /(SME+Enterprise)*100)

p9 <- plot_ly(data=by_value,x=~month,y=~cumulative_SME_perc,type="scatter",mode="lines+markers",name="Cumulative") %>%
  add_trace(y=~monthly_SME_perc,name="Monthly") %>%
  layout(title="Percentage of contract value awarded to SMEs",
         yaxis = list(title="Percentage of contract value awarded to SME"))
```  
`r p8`  
<br />  
`r p9`  

## Contract values over time  

Coming soon! 

```{r, echo = FALSE}
#c <- contracts %>% 
#  arrange(Publish.Date) %>%
#  select(Value, Publish.Date)

#c$trend <- loess(Value ~ seq_along(c$Publish.Date), data = c, span = 0.5)$fitted

#plot_ly(data = c, x =~ Publish.Date, y =~ Value, type = "scatter", mode = "markers") %>%
#  add_trace(y =~ trend, mode = "lines") %>%
#  layout(yaxis = list(type = "log"))
```
## Top buyers publishing  
## Top buyers by contracts published  
## New agencies this month  

