---
title: "Specialist Pricing"
output:
  html_document:
    theme: cerulean
  pdf_document: default
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# if rendered in a new R session, load the session values
if (!exists("buyers")) {
  source("R\\utilities.r")
  source("R\\plotly_functions.r")
  source("R\\assorted_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  library(stringr)
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),rel_path_data,substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
}

#reference data
thresholds = tibble(
  aoe = c( 
    "Agile delivery and Governance"    ,        "Software engineering and Development",    
    "User research and Design"          ,       "Change and Transformation"            ,   
    "Cyber security"                     ,      "Strategy and Policy"                   ,  
    "Marketing, Communications and Engagement", "Data science"                           , 
    "Emerging technologies" ,                   "Content and Publishing"                  ,
    "Support and Operations" ,                  "Training, Learning and Development"      ),
  maxprice = c(
    1767, 1760,
    1906, 2420,
    2200, 2491,
    2068, 1894,
    1925, 1472,
    1384, 2420
  )
)
```
# Pricing progress
```{r echo=FALSE}

# Note there are many 'prices' without matching domain, or entered by recruiters - so this is not entirely accurate
sellers$has_pricing <- apply(select(sellers,starts_with("maxPrice")),1,function(x) sum(x,na.rm=TRUE)) > 0
# remove the recruiters who are not also consultants
s <- sellers %>% filter(!is_recruiter|(is_recruiter&no_of_case_studies>0))
no_recruiters <- nrow(sellers) - nrow(s)
no_legacy     <- sellers %>% filter(nchar(legacy_aoe)>1,!dmp_framework) %>% nrow
product_only  <- sum(sellers$product_only)
total_sellers <- nrow(sellers) - no_recruiters - no_legacy - product_only
# need to filter out pricing that doesn't match a nominated category
# List of all anticipated pricings, with any prices provided to date
s2 <- s %>% 
  mutate(aoe_all = paste(assessed_aoe,unassessed_aoe,sep="|")) %>%
  select(aoe_all,code,name,starts_with("maxPrice")) %>%
  gather(starts_with("maxPrice"),key="aoe",value="price") %>%
  arrange(code)

s2$aoe <- gsub("maxPrice\\.","",s2$aoe)
s2$aoe <-  gsub("_"," ",s2$aoe)
matches <- nzchar(str_match(s2$aoe_all,s2$aoe),keepNA = TRUE)
matches[is.na(matches)] <- FALSE
s2$match <- matches
# also filter out the change, training and transformation prices - we don't care about them any more
s2 <- s2 %>% filter(s2$match,!grepl("Training and Transformation",aoe))
# List of sellers and their current pricing status
s3 <- s2 %>%
  select(code,name,aoe,price) %>%
  arrange(code) %>%
  group_by(code,name) %>%
  summarise(complete_pricing = !is.na(sum(price)),
            partial_pricing  = sum(price,na.rm=TRUE)>0)

s_partials <- s3 %>% filter(!complete_pricing,partial_pricing)
# s4 detailed list of all pricing values provided to date - filtered to remove any spurious prices
s4 <- s2 %>% filter(!is.na(price)) %>%
  left_join(thresholds, by="aoe") %>%
  select(-match,-aoe_all) %>%
  mutate(over_max = (price > maxprice))

x <- s2 %>% filter(code %in% s_partials$code)

## look at the edit requests - some edits are for sellers that already have pricing, so discount them
apps$has_pricing    <- apply(select(apps,starts_with("maxPrice")),1,function(x) sum(x,na.rm=TRUE)) > 0
s_with_pricing <- sellers %>% filter(has_pricing) %>% select(code)
a_submitted <- apps %>% 
  filter(type=="edit",status=="submitted",has_pricing) %>%
  filter(!code %in% s_with_pricing$code)
a_saved <- apps %>% 
  filter(type=="edit",status=="reverted",has_pricing) %>%
  filter(!code %in% s_with_pricing$code)

no_remaining <- total_sellers - sum(s3$partial_pricing) - nrow(a_submitted) - nrow(a_saved)
```
### Seller breakdown    
  
.|Count|%  
--|--|--   
Number of sellers in catalogue | `r nrow(sellers)` | 100%  
Number of recruiters (pricing not required) | `r no_recruiters` | `r round(no_recruiters/nrow(sellers)*100)`%  
DSPP sellers (inactive) | `r no_legacy` | `r round(no_legacy/nrow(sellers)*100)`%  
Product only sellers| `r sum(sellers$product_only)` | `r round(sum(sellers$product_only)/nrow(sellers)*100)`%  
Sellers requiring prices | `r total_sellers` | `r round(total_sellers/nrow(sellers)*100)`%  

### Current progress    
.|Count|%  
--|--|--   
Sellers who require pricing | `r total_sellers` | 100%  
Sellers with some pricing | `r sum(s3$partial_pricing)` | `r round(sum(s3$partial_pricing)/total_sellers*100)`%
Sellers with full pricing | (`r sum(s3$complete_pricing)`) | (`r round(sum(s3$complete_pricing)/total_sellers*100)`%)  
Sellers with edit requests submitted | `r nrow(a_submitted)` | `r round(nrow(a_submitted)/total_sellers*100)`%   
Sellers with edit requests saved | `r nrow(a_saved)` | `r round(nrow(a_saved)/total_sellers*100)`%   
Remaining | `r no_remaining` | `r round(no_remaining/total_sellers*100)`%  
~edit requests are counted only if they contain pricing while the seller profile does not.  
  
```{r, echo=FALSE}
s_with_over_max <- s4 %>% 
  group_by(code,name,over_max) %>% 
  summarise(count = n()) %>% 
  filter(over_max)
s_with_under_max <- s4 %>%
  group_by(code,name,over_max) %>% 
  summarise(count = n()) %>% 
  filter(!over_max)
s_with_both <- sum(s_with_over_max$code %in% s_with_under_max$code)
```
### By domain  

.|Count|%  
--|--|--   
Total prices anticipated | `r nrow(s2)` | 100%
Total prices provided | `r nrow(s4)` | `r round(nrow(s4)/nrow(s2)*100)`%  
Prices provided over maxprice | `r sum(s4$over_max)` | `r round(sum(s4$over_max)/nrow(s4)*100)`%  
  
### For sellers with pricing  

.|Count|%  
--|--|--   
Sellers with any prices over max | `r nrow(s_with_over_max)` | `r round(nrow(s_with_over_max)/total_sellers*100)`%  
Sellers with any prices under max | `r nrow(s_with_under_max)` | `r round(nrow(s_with_under_max)/total_sellers*100)`%  
Sellers with some (not all) prices over max | `r s_with_both` | `r round(s_with_both/total_sellers*100)`%  
Total | `r nrow(s_with_over_max)` + `r nrow(s_with_under_max)` - `r s_with_both` = `r nrow(s_with_over_max) + nrow(s_with_under_max) - s_with_both` |    



## Price ranges  

### All prices received  
```{r, echo=FALSE}  
x <- s4$price
hist(x,breaks=80,main="Histogram of all prices provided",ylab="count",xlab="$ (incl GST)")  
```  

### By area of expertise  
  
```{r, echo=FALSE}  
for (aoe_ in unique(s4$aoe)) {
  x <- s4 %>% filter(aoe == aoe_)
  x <- x$price
  hist(x,breaks=40,main=paste(aoe_,"prices"),ylab="count",xlab="$ (incl GST)")
  maxp <- thresholds[which(aoe_ == thresholds$aoe),]$maxprice
  cat(paste0("Threshold = $",maxp,"  \n"))
}
```  
  
  
---  



# Reference information for the pricing thresholds

Pricing information based on all responses to seller briefs from 01-04-2017 to 31-03-2018 (inclusive).

```{r, echo=FALSE}
start_date <- as.Date("2017-04-01")
end_date   <- as.Date("2018-03-31")
# dataset to use
br <- briefResponses %>%
    filter(type == "Specialist") %>% # only responses to all specialist roles
    filter(applicationDate >= start_date, applicationDate <= end_date) %>% # within date range
    filter(dayRate < 4000) # exclude extreme values
```

There were `r dim(br)[1]` specialist responses received in that time.  

### Count of bids received for each area of expertise  
```{r, echo=FALSE}
b <- br %>%  
    group_by(areaOfExpertise) %>%
    summarise(responses = n(),
              `90 percentile` = quantile(dayRate,0.9),
              `95 percentile` = quantile(dayRate,0.95),
              `10 percentile` = quantile(dayRate,0.1),
              `5 percentile` = quantile(dayRate,0.05))
b <- b %>% add_row(areaOfExpertise = "ALL",
              responses       = sum(.$responses),
              `90 percentile` = quantile(br$dayRate,0.9),
              `95 percentile` = quantile(br$dayRate,0.95),
              `10 percentile` = quantile(br$dayRate,0.1),
              `5 percentile` = quantile(br$dayRate,0.05))
b[which(b$areaOfExpertise=="Emerging technologies"),3:6] <- 
    b[which(b$areaOfExpertise=="ALL"),3:6]
kable(b)
```   
  
### Day rates bid for specialist briefs by service
```{r, echo=FALSE}
#plot_ly(x=~dayRate, color=~areaOfExpertise,data=br,type="box",
#        width=1000,height=500,colors=brewer.pal(11,"Paired")) %>% 
#  layout(showlegend=F,autosize=F,margin=list(l=200))



plot_ly(data = br,x=~dayRate,y=~areaOfExpertise,
        type="scatter",
        #color=~overThreshold,
        #colors=c("blue","red"),
        mode="markers",
        marker = list(size=2)) %>%
  layout(showlegend=F,autosize=F,
         margin=list(l=200),
         title="Day Rates Bid")

```  

## Seller activity versus pricing  
Check how active sellers are, versus whether they've provided pricing yet  

```{r, echo=FALSE}
br <- briefResponses %>%
  select(supplierId,supplierName,id,published,applicationDate) %>%
  left_join(s3,by=c("supplierId"="code")) %>%
  mutate(week = floor_date(applicationDate,unit="weeks"))
# recruiters are not in the s3 dataframe, but don't need pricing
br[is.na(br$partial_pricing),]$partial_pricing <- TRUE
br <- br %>%
  group_by(week,partial_pricing) %>%
  summarise(count=n())
x <- summary(br[!br$partial_pricing,]$count)
br$partial_pricing <- as.factor(br$partial_pricing)
plot_ly(data=br,x=~week,y=~count,color=~partial_pricing,type="scatter",mode="markers+lines") %>% 
  layout(title="Brief responses per week for sellers with/without pricing")

```
