---
title: "Seller extract"
output:
  word_document
params:
  seller: 
    label: seller
    value: "Seller name"
    input: textInput
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,include=FALSE}
# if rendered in a new R session, load the session values
if (!exists("buyers")) {
  source("utilities.r")
  source("plotly_functions.r")
  source("assorted_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  #library(kableExtra)
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),"\\..",rel_path_data,substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
}
```
  

```{r, echo=FALSE}
# lookup seller
s <- sellers %>% filter(grepl(params$seller,name,ignore.case = TRUE))
if (dim(s)[1] > 1) {
  print("**Warning, seller is ambiguous. Choose one the following and rerun**")
  print(s$name)
}
s_abn <- s[1,"abn"]
s_id  <- s[1,"code"]
s_summary <- tibble(parameter = names(s),value = as.character(s[1,]))
```  
**Date of extract: `r date_to_string(Sys.Date())`**    
  
#`r s$name`, (ABN: `r s_abn`)  

## Seller details  
https://marketplace.service.gov.au/supplier/`r s_id`    
  
`r kable(s_summary)`  

---  

## Reported Contracts  
From Austender, which reports contracts awarded to non corporate commonwealth entities over $10K. Recent contracts may not be reported yet.  


```{r,echo=FALSE}
c <- contracts %>% filter(Supplier.ABN==s_abn) %>%
  select(Agency,CN.ID,Description,Start.Date,End.Date,Value) %>%
  arrange(Start.Date)
```
Total contracts = `r dim(c)[1]`  
Total value     = `r format_mil(sum(c$Value))`    
`r kable(c)`

---  

## Applications for briefs  
```{r, echo=FALSE}  
br <- briefResponses %>% filter(supplierId==s_id) %>%
  select(agencyName, title, openTo, type, published) %>%
  arrange(published)
```
Total applications for briefs = `r dim(br)[1]`  
`r kable(br)`  

  

