---
title: "Snapshot"
output:
  html_document:
    theme: journal
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,include=FALSE}
# if rendered in a new R session, load the session values
if (!exists("buyers")) {
  source("utilities.r")
  source("plotly_functions.r")
  source("assorted_functions.r")
  library(knitr)
  library(tidyr)
  library(RColorBrewer)
  library(dplyr)
  library(plotly)
  library(httr)
  ## load the files - ensure these files are already produced
  load(paste0(getwd(),"\\..",rel_path_data,substr(as.POSIXct(Sys.time()),1,10),".Rdata"))
}

# filter inactive buyers
buyers <- buyers[buyers$active==TRUE,]
buyers <- buyers[!grepl("\\+",buyers$email_address),]
# need a better function to determine close date

```
**run date: `r date_to_string(Sys.Date())`**  
  

---   

## Current briefs  

```{r, echo = FALSE}
### Open to selecteds  
#`r kables(invited_sellers(briefs,briefResponses))`  
```
  
### Currently Live  
```{r echo=FALSE}
bl <- briefs[briefs$status=="live",] # 
#bl$closes <- string_to_date(bl$published) + 7
#bl[bl$duration=="2 weeks","closes"] <- bl[bl$duration=="2 weeks","closes"] + 7.5
bl$closes <- bl$close
bl <- bl[,c("id","title","organisation","openTo","type","closes")]
b_apps <- briefResponses[!is.na(briefResponses$applicationId),] %>% 
  group_by(id) %>% 
  summarise(applications = length(id))
bl <- merge(bl,b_apps,by.x="id",by.y="id",all.x=TRUE) %>% arrange(closes)
bl[is.na(bl$applications),]$applications <- 0
#bl$closes     <- date_to_string(bl$closes)
#bl <- bl[order(bl$closes),]
```
  
Number of open briefs: `r dim(bl)[1]`  

`r kables(bl)`  
  
### Recently Closed  
```{r echo=FALSE}
#brc <- briefs[briefs$status=="closed"
#                     &(Sys.Date() - string_to_date(briefs$published)) < 30,] 
#brc$brief_closed <- string_to_date(brc$published) + 7
#brc[brc$duration=="2 weeks","brief_closed"] <- 
#                    brc[brc$duration=="2 weeks","brief_closed"] + 7.5
#brc$brief_closed <- brc$close
#brc <- brc[,c("id","title","organisation","openTo","type","published","brief_closed")]
#brc <- merge(brc,b_apps,by.x="id",by.y="id",all.x=TRUE)
#brc <- brc[order(brc$brief_closed,decreasing=TRUE),]
#brc$brief_closed <- date_to_string(brc$brief_closed)
#brc[is.na(brc$applications),]$applications <- 0

brc <- briefs %>%
  filter(status=="closed",Sys.Date() - string_to_date(published) < 30) %>%
  mutate(brief_closed = close) %>%
  select(id,title,organisation,openTo,type,published,brief_closed) %>%
  left_join(b_apps,by="id") %>%
  replace_na(list(applications=0))
```
`r kables(brc)`  
  
---  


## Briefs  
  
### Briefs open by date  
```{r,echo=FALSE}
count_open_briefs <- function(date,b) {
  return(sum((b$published <= date)&(b$close >= date)))
}
count_open_briefs2 <- function(date,b) {
  return(sum(b$published <= date) - sum(b$close < date))
}

dsl <- datesSinceLaunch()$dates
x <- sapply(dsl,count_open_briefs,b=briefs)
y <- briefs %>% group_by(published) %>% summarise(count=n())

plot_ly(x=dsl,y=~x,mode="lines",type="scatter",name="briefs open") %>% 
  add_trace(data=y,x=~published,y=~count,type="bar",name="briefs published",inherit=FALSE) %>%
  layout(title="Brief publishing",
         xaxis=list(title=""),
         yaxis=list(title="Count"))
```
  
### Agencies publishing by month  
```{r,echo=FALSE}
b <- briefs %>% 
  mutate(month = floor_date(published,unit="month")) %>% 
  group_by(month) %>%
  summarise(count=length(unique(agencyName))) %>%
  filter(month < floor_date(Sys.Date(),unit="month")) # exclude the current month
plot_ly(data=b,x=~month,y=~count,type="scatter",mode="lines+markers") %>%
  layout(title="Number of agencies publishing per month",
         yaxis=list(range=c(0,40)))
```
  
### Published briefs
```{r echo=FALSE}
#b <- briefs[!duplicated(briefs[,c("id"),]),]
b <- briefs
b <- b[b$id > 0,]
b_counts <- b %>% 
  group_by(category) %>%
  summarise(count = length(category))
b_counts$percentage <- round(b_counts$count/sum(b_counts$count)*100)
b_counts <- rbind(b_counts,c("Total",sum(b_counts$count),100))
```
`r kables(b_counts)`  
  
#### Number of Briefs Published Since Launch  
`r plot_briefs_published_by_type(briefs)`  
  
### Responses for Specialist Briefs  
```{r, echo=FALSE}
# violin plots by area of expertise
vals <- briefResponses %>% filter(!is.na(areaOfExpertise),
                                  !is.na(dayRate),
                                  dayRate > 400,dayRate < 4000)
add_density_plot <- function(p,bR,i,color) {
#  dens <- density(bR$dayRate,na.rm=TRUE,from=(min(bR$dayRate)-100))
  dens <- density(bR$dayRate,na.rm=TRUE)
  up   <- dens$y/max(dens$y) + (i*3 + 2)
  down <- (i*3 + 2) - dens$y/max(dens$y)
  n    <- bR[1,"areaOfExpertise"]
  p <- p %>% add_lines(x=dens$x,y=up,mode="line",color=color,name=n)
  p %>% add_lines(x=dens$x,y=down,mode="line",color=color,
                  fill = "tonexty",
                  showlegend=FALSE)
}

resps <- vals %>% group_by(areaOfExpertise) %>% 
  summarise(number_of_briefs = length(unique(id)), 
            number_of_responses = length(applicationId))

areas <- data.frame(areaOfExpertise=unique(vals$areaOfExpertise))
doms  <- data.frame(areaOfExpertise= domains)
resps <- merge(resps,doms,all.y=TRUE)
resps[is.na(resps)] <- 0
names(resps) <- c("Area of Expertise","Number of Briefs","Number of Responses")
all_dens <- density(vals$dayRate,na.rm=TRUE)
up       <- all_dens$y/max(all_dens$y) + 2
down     <- 2 - all_dens$y/max(all_dens$y)
p <- plot_ly(x=all_dens$x,y=up,
             type="scatter",
             mode="lines",
             name="All areas",
             color=brewer.pal(10,"Paired")[1])
p <- p %>% add_lines(x=all_dens$x,y=down,mode="line",
                     color=brewer.pal(12,"Paired")[1],
                     fill = "tonexty",
                     showlegend=FALSE)
areas <- as.character(resps[resps$`Number of Responses`>10,]$`Area of Expertise`)
for (i in 1:length(areas)) {
#  if (resps[resps$areaOfExpertise==areas[i],]$number_of_responses > 9) {
    p <- add_density_plot(p,
                          vals[vals$areaOfExpertise==areas[i],],
                          i,
                          brewer.pal(12,"Paired")[i+1])
#  }
}
p <- p %>% layout(xaxis=list(range=c(0,3000),
                             title="Day Rate ($ incl GST)", showaxis=FALSE),
                  yaxis=list(showticklabels=FALSE,
                             showgrid=FALSE,
                             showline=FALSE),
                  #legend = list(orientation = 'h')
                  showlegend=FALSE)
                   #title="Distribution of Day Rates for Specialist Briefs")
# annotations
y_s   <- c(2,1:length(areas) * 3 + 2) + 0.6
areas <- c("All areas",areas)
p <- p %>% add_annotations(  
  x = rep(2500,length(areas)),
  y = y_s,
  showarrow=FALSE,
  text = areas
)

```  
  
  
#### Number of Specialist Briefs and Responses Since Launch  
`r kables(resps)`  
<br>  
  
#### Bids for Open to All Specialist Briefs  
`r plot_bids_by_specialist_brief(briefResponses)`  
<br>  
  
#### Distribution of bids for Specialist Briefs by domain  
`r p`
<br>   

#### Applications per brief for Open to All  
<br>  

```{r,echo=FALSE}
specialists <- briefResponses %>% 
  filter(openTo=="All", type=="Specialist") %>% 
  filter(published > Sys.Date() - 365) %>%
  group_by(id) %>% 
  summarise(responses = sum(attachments)) %>%
  filter(!is.na(responses))
outcomes <- briefResponses %>% 
  filter(openTo=="All", type=="Outcome") %>% 
  filter(published > Sys.Date() - 365) %>%
  group_by(id) %>% 
  summarise(responses = n()) %>%
  filter(!is.na(responses))
dens_s <- density(specialists$responses,from=0)
dens_s$y <- dens_s$y / (max(dens_s$y)) * 45
dens_o <- density(outcomes$responses,from=0)
dens_o$y <- dens_o$y / (max(dens_o$y)) * 45
plot_ly(alpha = 0.6) %>%
  add_histogram(x=specialists$responses,type="specialists",name="histogram") %>%
  add_histogram(x=outcomes$responses,type="outcomes",name="histogram") %>%
  add_lines(x=dens_s$x,y=dens_s$y,name="Specialists") %>%
  add_lines(x=dens_o$x,y=dens_o$y,name="Outcomes") %>%
  layout(
    barmode="overlay",
    title="Seller Responses per Brief",
    xaxis=list(title="Applications"),
    yaxis=list(title="Count"))
  

```
<br>  
Summary figures:  

Type | Min. | 1st Quartile | Median | Mean | 3rd Quartile | Max.  
--|--|--|--|--|--|--  
Specialists | `r paste(paste(summary(specialists$responses),collapse = " | "),"  ")`  
Outcomes | `r paste(paste(summary(outcomes$responses),collapse = " | "),"  ")`  

      
### Publishing Agencies  
```{r echo=FALSE}
b <- briefs[briefs$id > 0,c("id","organisation","buyerEmail","email_domain")]
b <- b[!duplicated(b$id),]
#b$domain <- matrix(unlist(strsplit(b$buyerEmail,"@")),ncol=2,byrow=TRUE)[,2]
counts <- b %>% group_by(email_domain) %>%
  summarise(count = length(id), agency = organisation[1]) %>%
  arrange(desc(count))
counts$email_domain <- factor(counts$email_domain,levels=counts$email_domain) 
#plot_ly(counts,x=~count,y=~domain,type="bar",orientation='h') %>%
#  layout(yaxis = list(title = #'',tickfont=list(color="darkblue"),font='arial',size=12),
#         xaxis = list(title = ''),
#         autosize=F,
#         width=1000,
#         height=1000,
#         margin = list(l = 200, r = 10, t = 10, b = 20))

```
   
   
**Total publishing agencies `r dim(counts)[1]`**    
  
  
`r kables(counts)`   

---  
  
## Sellers  
  
### Seller Activity  
`#r plot_application_counts_by_seller(briefResponses)`  
<br>  
```{r, echo=FALSE}
sellers$applied <- sellers$code %in% briefResponses$supplierId
sellers$status  <- "registered"
sellers[nchar(sellers$assessed_aoe) > 1,]$status  <- "active"
sellers[nchar(sellers$legacy_aoe) > 1 & sellers$dmp_framework,]$status  <- "upgraded"
sellers[!sellers$dmp_framework,]$status <- "legacy"
counts <- sellers %>% group_by(status, applied) %>% summarise(count =  length(applied))
as        <- sum(counts[counts$status %in% c("active","upgraded"),"count"])
asApplied <- sum(counts[counts$status %in% c("active","upgraded")&counts$applied,"count"])
asperc <- round(asApplied/(as)*100)
```
| | |
--|--
Total number of active sellers                 | `r as`  
Active sellers who've applied for any briefs   | `r asApplied`  
% active sellers who've applied for any brief  | `r asperc`  
Registered sellers | `r sum(counts[counts$status=="registered","count"])`  

  
```{r echo=FALSE}
a <- apps[apps$type %in% c("new","upgrade")&apps$status=="submitted",]
a$days_since_sub <- as.numeric(Sys.Date() - a$submitted_at_date)

dsppSU <- with(apps,sum(type %in% c("new","upgrade")&apps$status=="submitted"&type=="upgrade"))
dsppST <- with(apps,sum(type %in% c("new","upgrade")&apps$status%in%c("saved","reverted")&type=="upgrade"))
dsppNS <- sum(is.na(sellers$application_id)&!sellers$dmp_framework)
dsppAP <- with(sellers,sum(nchar(legacy_aoe)>1&dmp_framework))
dmpPROD <- with(sellers,sum(nchar(legacy_aoe)<2&product_only))
dmpASS  <- with(sellers,sum(dmp_framework&nchar(assessed_aoe)>1&nchar(legacy_aoe)<2&!product_only))
dmpREG <- with(sellers,sum(dmp_framework&nchar(assessed_aoe)<2&nchar(legacy_aoe)<2&!product_only))
totes <- dsppAP+dsppNS+dsppST+dsppSU + dmpASS+dmpPROD+dmpREG
```

### Seller counts  
  
```{r,echo=FALSE}
s <- sellers %>% 
  group_by(number_of_employees,joined) %>% 
  arrange(joined) %>% 
  summarise(count = length(number_of_employees)) %>% 
  mutate(csum = cumsum(count))
s$number_of_employees <- factor(s$number_of_employees)
p <- plot_ly(s,x=~joined,y=~csum,type="scatter",mode="lines",
             color=~number_of_employees,
             colors=brewer.pal(9,"Set1")) %>%
      layout(title="Number of registered sellers over time",
           xaxis=list(title=""),
           yaxis=list(title="Count"))
p
s <- sellers %>%
  arrange(joined)
s$csum <- 1:dim(s)[1]
plot_ly(s,x=~joined,y=~csum,type="scatter",mode="lines") %>%
      layout(title="Number of registered sellers over time",
           xaxis=list(title=""),
           yaxis=list(title="Count"))
lm <- lm(csum~joined,data=s[444:dim(s)[1],])
#s <- sellers %>% 
  #filter(dmp_framework) %>% 
#  arrange(joined) %>% 
#  mutate(number = 1:length(joined))
#plot_ly(s,x=~joined,y=~number,type="scatter",mode="lines") %>%
#    layout(title="Number of registered sellers over time",
#           xaxis=list(title=""),
#           yaxis=list(title="Count"))


```
  
Shows all sellers who are *currently on the Marketplace* based on when they were added as a seller, and split by the nunber of employees each has. Note that this shows registered sellers, not all of these will be active.  
  
  
### Seller composition  
  
#### Category counts  
```{r,echo=FALSE}
aoes <- sapply(domains,function(x,s) {
        sum(grepl(x,s$assessed_aoe))
       },s = sellers)
aoes_un <- sapply(domains,function(x,s) {
        sum(grepl(x,s$unassessed_aoe))
       },s = sellers)
aoe_df <- data.frame(domains,aoes,aoes_un)
#aoe_df <- aoe_df %>% arrange(aoes)
plot_ly(aoe_df,x=~domains,y=~aoes,name="Assessed",type="bar") %>%
  add_trace(y=~aoes_un,name="Not yet assessed") %>%
  layout(yaxis = list(title = 'Seller count by Area of Expertise'),
         xaxis = list(title = ""), 
         barmode = 'stack')
```

  
#### Registered Sellers  
  
| buyer type and status | number |  
|--|--|  
| DSPP not started | `r dsppNS` |  
| DSPP started | `r dsppST` |  
| DSPP submitted | `r dsppSU` |  
| DSPP approved | `r dsppAP` |  
| * DSPP subtotal * | * `r sum(dsppNS,dsppST,dsppSU+dsppAP) ` * |
| DMP registered | `r dmpREG` |  
| DMP active | `r dmpASS` |  
| DMP product only | `r dmpPROD` |  
| * DMP (new sellers) subtotal * | * `r sum(dmpREG,dmpASS,dmpPROD) ` * |
| Total of the above | `r totes` |  
| Cross check (should equal the 'Total of the above') | `r dim(sellers)[1]` |  
  
#### Applications  
  
```{r echo=FALSE}
ipREV <- with(apps,sum(status=="reverted"))
ipABN <- with(apps,sum(status=="saved"&!is.na(abn)&grepl("\\d",abn)))
ipNO_ABN <- with(apps,sum(is.na(abn)) + sum(!is.na(abn)&!grepl("\\d",abn)))
rejected <- with(apps,sum(status=="assessment_rejected"))
```  
| Applicant status | number |  
|--|--|  
| Submitted, awaiting assessment | `r sum(a$type=="new")`|  
| In progress - reverted | `r ipREV` |  
| In progress - not yet submitted, has entered ABN | `r ipABN`|  
| In progress - not yet submitted, has not entered ABN | `r ipNO_ABN`|  
| Rejected outright | `r rejected` |  
  

---  

## Applications  
  
---  

## Operations  
  
### Domain assessments  

  
### New seller assessments

```{r echo=FALSE}
#a[is.na(a$days_since_sub),]$days_since_sub <- 0
xmax <- max(a[a$type %in% c("new","upgrade"),]$days_since_sub)
xbins <- list(start=1,end=xmax,size=1)
p <- plot_ly(x = a[a$type=="new","days_since_sub"],type="histogram",name="new",xbins=xbins)
#if (length(a[a$type=="new","days_since_sub"])>0) {
##  p <- add_histogram(p,x = a[a$type=="new","days_since_sub"],name="new",xbins=xbins) 
#}
if (length(a[a$type=="upgrade","days_since_sub"]) > 0) {
  p <- add_histogram(p,a[a$type=="upgrade","days_since_sub"],name="upgrade",xbins=xbins)
} 
p <- layout(p,barmode="overlay",
         title = 'Unassessed Application or Upgrades',
           xaxis = list(title = "days since submission",
                        showgrid = TRUE),
           yaxis = list(title = "number of submissions",
                        showgrid = TRUE))
p
```

     
  
| Total of | Count |
|---|---|  
| New applications not yet assessed | `r sum(a$type=="new")` |  
| Upgrades not yet assessed | `r sum(a$type=="upgrade")` |  

### Seller edits
```{r echo=FALSE}
a <- apps[apps$type=="edit"&apps$status=="submitted",]
a$days_since_sub <- as.numeric(Sys.Date() - a$submitted_at_date)
plot_ly(x=a$days_since_sub,type="histogram",nbinsx=20) %>% 
  layout(barmode="overlay",
         title = 'Unassessed Edit Requests',
           xaxis = list(title = "days since submission",
                        showgrid = TRUE),
           yaxis = list(title = "number of submissions",
                        showgrid = TRUE))

```
  
---  
  
## Buyers  
  
### Buyer and Agency Counts  
  
  
```{r echo=FALSE}
plot_buyer_counts_2(buyers) 
agencies <- buyers %>%
                group_by(category) %>% 
                summarise(count=length(category))  
agencies$percent <- round(agencies$count/sum(agencies$count) * 100,digits=1)
agencies <- rbind(agencies,c("Total buyers",sum(agencies$count),100.0))
```  
  
* Active buyers have signed in or registered in the past 90 days

  
### Buyer Counts by Agency Category  
`r kables(agencies)`  
  
`r plot_buyer_count_by_agency_category(buyers)`  
    

  
### Agency Counts by Category  
```{r echo=FALSE}
agencies_counts <- buyers %>%
                group_by(email_domain,category) %>% 
                summarise(count=length(email_domain)) %>%
                group_by(category) %>%
                summarise(count = length(category))
agencies_counts$percent <- round(agencies_counts$count/sum(agencies_counts$count) * 100,digits=1)
agencies_counts <- rbind(agencies_counts,c("Total agencies",sum(agencies_counts$count),100.0))

```
`r kables(agencies_counts)`     
  
`r plot_agency_count_by_category(buyers)`    

---   



