---
title: "Marketplace Transactions"
output:
  html_document:
    theme: journal
    toc: true
    toc_float: true
  pdf_document: default
  word_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#source("utilities.r")
#library(knitr)
#library(tidyr)
loglabs <- c("10K","100K","1M","10M")
#load(paste(substr(as.POSIXct(Sys.time()),1,10),".Rdata",sep=""))
# load Austender data
  # version to extract data from a Google Sheet
  #aus_ws <- gs_title("Austender Data")
  #sonAll <-  aus_ws %>% gs_read(ws="All")
  # and regularise it
  #sonAll$StartDate <- string_to_date(sonAll$Start_Date, format = "%d-%b-%y")
  #sonAll$EndDate   <- string_to_date(sonAll$End_Date, format = "%d-%b-%y")
  #sonAll$PublishDate   <- string_to_date(sonAll$Publish_Date, format = "%d-%b-%y")
sonAll <- contracts
sonAll <- sonAll[order(sonAll$Publish.Date),]
sonAll$cumulative_value <- cumsum(sonAll$Value)
sonAll$cumulative_contracts <- 1:dim(sonAll)[1]
sonAll$LogValue <- log10(sonAll$Value)
# add further seller details
s <- sellers
#s$abn <- as.numeric(s$abn)

# contracts is now merged with seller information in an earlier function 
#contracts <- left_join(sonAll,s,by=c("Supplier.ABN" = "abn"))


# calcs the coordinates for an AB line to show a linear regression
abline <- function(coef,start,end) {
  start = c(start,coef[1]+coef[2]*as.numeric(start))
  end   = c(end,coef[1]+coef[2]*as.numeric(end))
  return(data.frame(matrix(c(start,end),nrow=2,ncol=2,byrow=TRUE)))
}

latest_contract <- max(sonAll$Publish.Date)
sonDSP <- sonAll[sonAll$SON.ID=="SON3364729",]
sonDMP <- sonAll[sonAll$SON.ID=="SON3413842",]
```
**Run date: `r date_to_string(Sys.Date())`  
**Latest contract in report: `r date_to_string(latest_contract)`  



## Briefs Published by Date  
```{r, echo=FALSE}
b <- briefs
b <- b[order(b$published),]
b$cumulative_number <- 1:dim(b)[1]

b <- briefs %>%
  mutate(ncce=(entities=="NCCE")) %>%
  group_by(ncce,published) %>%
  arrange(published) %>%
  summarise(count = length(ncce)) %>%
  mutate(cumulative_number = cumsum(count)) %>%
  arrange(published)

```



### Briefs and Contracts by Publishing Date    
<br>  
```{r echo=FALSE}
# linear model before and including 23/May
lm1 <- lm(cumulative_contracts~Publish.Date,
          data=sonAll[sonAll$Publish.Date<as.Date("2017-05-24"),])
# linear model after 23/May
lm2 <- lm(cumulative_contracts~Publish.Date,
          data=sonAll[sonAll$Publish.Date>as.Date("2017-05-23"),])

# for briefs after 13/Apr
lm3 <- lm(cumulative_number~published,
          data=b[b$published>as.Date("2017-04-12")&b$ncce,])

# briefs in the last 30 days
lm4 <- lm(cumulative_number~published,
          data=b[b$published>(Sys.Date()-31)&b$ncce,])

points1 <- abline(lm1$coefficients,as.Date("2016-08-01"),as.Date("2017-09-30"))
points1$X1 <- as.Date(points1$X1,origin="1970-01-01")
points2 <- abline(lm2$coefficients,as.Date("2017-05-23"),Sys.Date())
points2$X1 <- as.Date(points2$X1,origin="1970-01-01")
points3 <- abline(lm3$coefficients,as.Date("2017-04-12"),Sys.Date())
points3$X1 <- as.Date(points3$X1,origin="1970-01-01")
points4 <- abline(lm4$coefficients,Sys.Date()-45,Sys.Date()+15)
points4$X1 <- as.Date(points4$X1,origin="1970-01-01")

p <- plot_ly(data=sonAll,x=~Publish.Date,y=~cumulative_contracts,
             mode="markers",type="scatter",name="contracts")
b_t <- b[b$ncce,]
p <- p %>% add_markers(data=b_t,x=~published,y=~cumulative_number,
                       mode="markers",type="scatter",name="NCCE briefs")
b_t <- b[!b$ncce,]
p <- p %>% add_markers(data=b_t,x=~published,y=~cumulative_number,
                       mode="markers",type="scatter",name="Non-NCCE briefs")
#p <- p %>% add_lines(data=points1,x=~X1,y=~X2,mode="line",name="trend line",
#                     line=list(dash='dot'))
p <- p %>% add_lines(data=points2,x=~X1,y=~X2,mode="line",name="trend line 1",
                     line=list(dash='dot'))
p <- p %>% add_lines(data=points3,x=~X1,y=~X2,mode="line",name="trend line 2",
                     line=list(dash='dot'))
p <- p %>% add_lines(data=points4,x=~X1,y=~X2,mode="line",name="last 30 days",
                     line=list(dash='dot'))
p %>%    layout(title = 'Number of Briefs and Contracts by Published Date',
                    xaxis = list(title = "Published Date",
                                 showgrid = TRUE),
                    yaxis = list(title = "Number",
                                 showgrid = TRUE))

```
<br />  
  
#### Contracts published per day:  

Briefs/Contracts published per day | |
--|--|--      
Contracts before 23/May/17 | <strong>`r round(lm1$coefficients[2],2)`</strong>    
Contracts after 23/May/17 (trend line 1) | <strong>`r round(lm2$coefficients[2],2)`</strong>  
NCCE briefs after 12/Apr/17 (trend line 2) | <strong>`r round(lm3$coefficients[2],2)`</strong>  
NCCE briefs (last 30 days) | <strong>`r round(lm4$coefficients[2],2)`</strong>  
  
### Contracts published in the last week  
```{r echo=FALSE}
last_contracts <- contracts %>%
  filter(Publish.Date > Sys.Date() - 8) %>%
  select(Agency, name, Supplier.ABN, Publish.Date, Value, sme_by_employees) %>%
  arrange(desc(Publish.Date))
```  
  
`r kables(last_contracts)`

### Contracts varied in the past week  
```{r echo=FALSE}
#rel_path_data()
old_contracts <- last_contracts_file()
contract_variations <- ready(old_contracts) %>%
  select(CN.ID,Value) %>%
  left_join(contracts,by="CN.ID") %>%
  mutate(Variation = Value.y - Value.x) %>%
  filter(Variation != 0) %>%
  select(Agency,name,Supplier.ABN,Publish.Date,Value.y,Variation)
names(contract_variations) <- qw("Agency Seller_Name ABN First_Published New_Value Variation")
#kables(contract_variations)
```  
Contracts that have been varied since `r str_match(old_contracts,"\\d{4}\\-\\d{2}\\-\\d{2}")[1,1]`  
Total variation of `r format_mil(sum(contract_variations$Variation),digits=3)`  
`r kables(contract_variations)`  
  
  
### Contract Values      
<br />    

Contracts | Number | Value  
--|--|--  
Total | `r dim(sonAll)[1]` | $`r round(sum(sonAll$Value)/1000000,2)`M |  
DSPP (SON3364729) | `r dim(sonDSP)[1]` | $`r round(sum(sonDSP$Value)/1000000,2)`M   
DMP (SON3413842) | `r dim(sonDMP)[1]` | $`r round(sum(sonDMP$Value)/1000000,2)`M  
<br />   


```{r echo=FALSE}
lm_con <- lm(cumulative_value~Publish.Date,
             data=sonAll[sonAll$Publish.Date>(Sys.Date()-90),])
line <- abline(lm_con$coefficients,Sys.Date()-90,Sys.Date()+15)
line$X1 <- as.Date(line$X1,origin="1970-01-01")
p <- plot_ly()
points <- subset(sonAll,SON.ID=="SON3413842")
# polynomial trend line
end_date   <- floor_date(Sys.Date(),"month")
start_date <- as.Date("2017-05-30")
months     <- interval(start_date,end_date) %/% months(1)
x_a        <- sonAll %>% filter(Publish.Date <= start_date)
x_a        <- sum(x_a$Value)
x_b        <- sonAll %>% filter(Publish.Date <= end_date)
x_b        <- sum(x_b$Value)
perc       <- 10^(log10(x_b/x_a)/months) - 1
perc       <- ceiling(perc*100)/100 # round to the nearest percentile
months     <- data.frame(months = seq(start_date,ceiling_date(Sys.Date(),"month"),by="months"))
months$count <- 0:(nrow(months)-1)
months$estimate <- x_a * (1+perc) ^ months$count
caption    <- paste0("growth at ",round(perc*100),"%")

p <- p %>% add_trace(data=points,x=~Publish.Date,y=~cumulative_value,
                     type="scatter",mode="markers",name="Marketplace Panel",
                     marker=list(size=5))
points <- subset(sonAll,SON.ID=="SON3364729")
p <- p %>% add_trace(data=points,x=~Publish.Date,y=~cumulative_value,
                     type="scatter",mode="markers",name="DSPP",
                     marker=list(size=5))
p <- p %>% add_lines(data=line,x=~X1,y=~X2,
                     name="past 90 days",
                    line=list(dash='dot'))
p <- p %>% add_lines(data=months,x=~months,y=~estimate,
                     name=caption,
                    line=list(dash='dot'))
p <- p %>% layout(title = 'Contracts published by Standing Offer',
                    xaxis = list(title = "Published Date",
                                 showgrid = TRUE),
                    yaxis = list(title = "Value ($)",
                                 showgrid = TRUE))
p
```
  
* Trend line shows contracts published in the past 90 days (since `r date_to_string(Sys.Date()-90)`)  
Rate = `r format_mil(lm_con$coefficients[2],order="K")` per day

<br>  
```{r, echo=FALSE}
# cumulative value by SME
cbd <- datesSinceLaunch()
# sums or counts all values up to and including the refdate
cumulation <- function(refdate,dates,values,type = "sum") {
  if (type=="sum") {
    return(round(sum(values[dates <= refdate])/1000000,2))
  }
  return(length(values[dates <= refdate]))
}
c <- contracts
cbd$cum_value  <- sapply(cbd$dates,cumulation,c$Publish.Date,c$Value)
cbd$cum_number <- sapply(cbd$dates,cumulation,c$Publish.Date,c$Value,"length")
c <- contracts[contracts$sme_by_employees,]
cbd$cum_sme_value  <- sapply(cbd$dates,cumulation,c$Publish.Date,c$Value)
cbd$cum_sme_number <- sapply(cbd$dates,cumulation,c$Publish.Date,c$Value,"length")
c <- contracts[!contracts$sme_by_employees,]
cbd$cum_e_value  <- sapply(cbd$dates,cumulation,c$Publish.Date,c$Value)
cbd$cum_e_number <- sapply(cbd$dates,cumulation,c$Publish.Date,c$Value,"length")

p <- plot_ly(cbd, x= ~dates, y = ~cum_value, name = 'SME', type='scatter',
        mode= 'non', fill = 'tozeroy', hoverinfo= 'text',
        text = ~paste(date_to_string(cbd$dates),
                       "<br />SME -<br />Contracts: ",cbd$cum_sme_number,
                      "<br />Value: $",cbd$cum_sme_value,
                       "M<br />Of total -<br />Contracts: ",
                       cbd$cum_number,
                       "<br />Value: $",cbd$cum_value,"M")) %>% 
  add_trace(y = ~cum_e_value, name='Enterprise', 
            text = ~paste(date_to_string(cbd$dates),
                       "<br />Enterprise -<br />Contracts: ",cbd$cum_e_number,
                      "<br />Value: $",cbd$cum_e_value,
                       "M<br />Of total -<br />Contracts: ",
                       cbd$cum_number,
                       "<br />Value: $",cbd$cum_value,"M")) %>% 
  layout(title = 'Cumulative value by seller type',
         xaxis = list(title = "Published Date",
                      showgrid = TRUE),
         yaxis = list(title = "Cumulative Contract Value ($M)",
                      showgrid = TRUE))


```
<br>  
```{r, echo=FALSE}
#lm <- lm(LogValue~Publish.Date,data=sonAll)
#tl <- abline(lm$coefficients,
#            min(datesSinceLaunch()$dates),
#            max(datesSinceLaunch()$dates))
plot_ly(data=sonAll,x=~Publish.Date,y=~LogValue,
        color=~SON.ID,colors=c("red","blue"),
        type="scatter",mode="markers") %>%
  #add_lines(data=tl,x=~X1,y=~X2,
  #                  name="trend line",
  #                  line=list(dash='dot')) %>% 
  layout(title = 'Value of Individual Contracts on Austender',
                    xaxis = list(title = "Published Date",
                                 showgrid = TRUE),
                    yaxis = list(title = "Value ($)",
                                 ticktext = loglabs,
                                 tickvals=c(4,5,6,7),
                                 showgrid = TRUE))

```
  
<br>   
```{r,echo=FALSE}
breaks <- 20
vals3364729 <- sonAll[sonAll$SON.ID=="SON3364729",]$LogValue
max_val_3364729 <- max(hist(vals3364729,breaks=breaks,plot=F)$counts)
dens3364729 <- density(vals3364729)
dens3364729$y <- dens3364729$y / max(dens3364729$y) * max_val_3364729
vals3413842 <- sonAll[sonAll$SON.ID=="SON3413842",]$LogValue
max_val_3413842 <- max(hist(vals3413842,breaks=breaks,plot=F)$counts)
dens3413842 <- density(vals3413842)
dens3413842$y <- dens3413842$y / max(dens3413842$y) * max_val_3413842
```
```{r, echo=FALSE}
p <- plot_ly(alpha = 0.6) %>%
  add_histogram(x = vals3364729,name="DSPP contracts",nbinsx=breaks) %>%
  add_histogram(x = vals3413842,name="DMP contracts",nbinsx=breaks) %>%
  add_lines(x=~dens3364729$x,y=dens3364729$y,name="DSPP smoothed") %>%
  add_lines(x=~dens3413842$x,y=dens3413842$y,name="DMP smoothed") %>%
  layout(barmode = "overlay",
         title = "Number of Contracts by Value",
         xaxis = list(title = "Value ($)",ticktext = loglabs, tickvals=c(4,5,6,7),
                                 showgrid = TRUE),
                    yaxis = list(title = "Count of contracts",
                                 showgrid = TRUE))
p

```

```{r, echo=FALSE}
pubRate <- round(lm3$coefficients[2],2)
dmpVals <- sonAll[sonAll$SON.ID=="SON3413842",]$Value
dspVals <- sonAll[sonAll$SON.ID=="SON3364729",]$Value
dmAv    <- round(mean(dmpVals)/1000,1)
dmMed   <- round(median(dmpVals)/1000,1)
dspAv   <- round(mean(dspVals)/1000,1)
dspMed  <- round(median(dspVals)/1000,1)
proj    <- round((dmAv*1000 * lm2$coefficients[2] * 365)/1000000)
```

Current Values for Digital Marketplace Panel  | Value     
--|--  
Current contract publishing rate | `r round(lm2$coefficients[2],2)` contracts per day  
Average value per contract       | $`r dmAv`K  
Median value per contract        | $`r dmMed`K  
Projected annual value           | $`r proj`M  
  
Values for the Digital Services Professionals Panel  | Value     
--|--  
Average value per contract       | $`r dspAv`K  
Median value per contract        | $`r dspMed`K  
  
  
  
## Sellers  
  
```{r, echo=FALSE}
#s <- sellers
#s$abn <- as.numeric(s$abn)
#contracts <- left_join(sonAll,s,by=c("Supplier_ABN" = "abn"))
sme_badge <- contracts %>% group_by(sme_badge) %>% summarise(value = sum(Value),count = length(sme_badge))
sme_badge$percent_value <- round(sme_badge$value / sum(sme_badge$value) * 100)
sme_badge$percent_count <- round(sme_badge$count / sum(sme_badge$count) * 100)
#sme_badge <- bind_rows(sme_badge,c("Total",sum(sme_badge$value),sum(sme_badge$count),sum(sme_badge$percent_value),sum(sme_badge$percent_count)))
sme_emps  <- contracts %>% group_by(sme_by_employees) %>% summarise(value = sum(Value), count = length(sme_by_employees))
sme_emps$percent_value <- round(sme_emps$value / sum(sme_emps$value) * 100)
sme_emps$percent_count <- round(sme_emps$count / sum(sme_emps$count) * 100)

```
### Contracts by Seller  
```{r, echo=FALSE}
cbs <- contracts %>% group_by(Supplier.ABN,name,sme_badge,is_recruiter,code) %>%
  summarise(number_contracts = length(Supplier.ABN),
            total = sum(Value),
            max=max(Value),
            min=min(Value),
            average=mean(Value)
            ) %>%
  arrange(desc(number_contracts))
app_counts <- briefResponses %>%
    group_by(supplierId, supplierName) %>%
    summarise(count = length(applicationDate)) %>%
    arrange(desc(count))
merged <- merge(cbs,app_counts,by.x="code",by.y="supplierId",all.y=TRUE) %>% 
  arrange(desc(count))
merged[is.na(merged)] <- 0
merged$index <- 1:dim(merged)[1]
merged$has_contracts <- merged$number_contracts > 0
plot_ly(data=merged,x=~index,y=~count,type="scatter",
        mode="markers",name="brief applications",marker=list(size=4),
        text = paste(merged$supplierName,
                     "<br />Applications: ",merged$count,
                     "<br />Contracts: ",merged$number_contracts
                       )) %>% 
  add_trace(x=~index,y=~number_contracts,mode="markers",name="contracts",
        text = paste(merged$supplierName,
                     "<br />Applications: ",merged$count,
                     "<br />Contracts: ",merged$number_contracts
                       )) %>% 
  layout(title="Counts of brief applications and reported contracts",
         xaxis=list(title=""),
         yaxis=list(range=c(0,40)))  
```
  
  
#### Top ten sellers by number of contracts    
`r kables(cbs[1:10,])`  
  
#### Contracts versus Brief Applications  
```{r, echo=FALSE}
x11 <- dim(sellers)[1]
x12 <- 100
x21 <- dim(app_counts)[1]
x22 <- round(x21 / x11 * 100)
x31 <- dim(cbs)[1]
x32 <- round(x31/x11 * 100)
x41 <- dim(sellers[sellers$dmp_framework&nchar(sellers$assessed_aoe) > 1,])[1]
x42 <- round(x41/x11 * 100)
```

Figure | number | percentage  
--|--|--    
Registered sellers in the Marketplace | `r x11` | `r x12`  
Sellers who've applied for briefs | `r x21` |  `r x22` 
Sellers awarded contracts | `r x31` |  `r x32`  
Number of active sellers | `r x41` | `r x42`   

### By SME Badge
`r kables(sme_badge)`  
  
### By Declared Number of Employees  
`r kables(sme_emps)`




